---
title: "COVIDMINDER Example Notebook: BOOTCAMP DEMO EDITS"
subtitle: "DAR/HACL Assignment 2 (Summer 2020)"
author: "Rufeng Ma"
date: "07/15/2020"
output: html_document
---

```{r setup, include=FALSE}
# Set the default CRAN repository
local({r <- getOption("repos")
       r["CRAN"] <- "http://cran.r-project.org" 
       options(repos=r)
})

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "../")
```


## INTRODUCTION

This R Notebook and its related R scripts provide a very basic introduction to the **COVIDMINDER** application, which you can find at:

* Public "production" version: https://covidminder.idea.rpi.edu/

The github repository for all the code required for this notebook, including a snapshot of the **COVIDMINDER** application, can be found at:

* https://github.com/TheRensselaerIDEA/COVID-Notebooks

The **COVIDMINDER** github repository can be found at:

* https://github.com/TheRensselaerIDEA/COVIDMINDER

**NOTE:** To contribute or submit to these repositories you must have a github.com ID and must send it to John Erickson at `erickj4@rpi.edu`. **PLEASE DO THIS IMMEDIATELY BEFORE READING ANY FURTHER!!**

## HACL ASSIGNMENT 2: CLONING A NOTEBOOK AND UPDATING THE REPOSITORY

In this assignment we're asking you to...

* clone the `COVID-Notebook` github repository...
* create a personal branch using git, and...
* make additions to the repository by creating a new, customized notebook. 

The instructions which follow explain how to accomplish this. 

**For DAR/HACL Summer 2020** you *must* be using RStudio Server on the IDEA Cluster. Instructions for accessing "The Cluster" appear at the end of this notebook. Don't forget to send your github ID to `erickj4@rpi.edu` 

The recommended procedure for cloning and using this repository is as follows:

* Access RStudio Server on the IDEA Cluster at http://lp01.idea.rpi.edu/rstudio-ose/
* Access the Linux shell on the IDEA Cluster by clicking the **Terminal** tab of RStudio Server. `cd` into your home (login) directory using: `cd ~`
    * Advanced users may use `ssh` to directly access the Linux shell
* `git clone https://github.com/TheRensselaerIDEA/COVID-Notebooks.git` in your `home` directory, resulting in a new directory `COVID-Notebooks`
* In the Linux shell, `cd` to `COVID-Notebooks/HACL-2020` 
* In the Linux shell, `git checkout -b hacl-yourrcs` where `yourrcs` is your RCS id
    * For example, if your RCS is `erickj4`, your new branch should be `hacl-erickj4`
    * It is critical that you include your RCS id in your branch id
* In the RStudio Server UI, navigate to the `COVID-Notebooks/HACL-2020` directory via the **Files** panel
    * Set this to be your R working directory (under "More" menu)
* **REQUIRED**: 
    * Make a <b>copy</b> of this notebook (ie `covid-notebook-hacl.Rmd`) file using a *new, original, descriptive* filename that **includes your RCS ID!**
    * Edit your new notebook; save
    * Use the RStudio `knit` comment to creat an HTML file; repeat as necessary
    * In Linux, `git add` each new file you want to add to the repository, esp. your new `Rmd` file and the `html` you created when you knitted
    * When you're ready, in Linux: `git commit -a -m "some comment"` where "some comment" is a useful comment describing your changes
    * Finally, `git push origin hacl-yourrcs` (where `hacl-yourrcs` is the branch you're working on) 
* **REQUIRED**: On github, select your branch (from the branch drop-down) and submit a pull request.

**REQUIRED:** For this assignment you will be asked to confirm the following in LMS:

    * The location of the github: https://github.com/TheRensselaerIDEA/COVID-Notebooks.git
    * Your github ID: VarunNair22
    * The name of your new branch: nairv2
    * The location where you committed your new (copied) notebook: https://github.com/TheRensselaerIDEA/COVID-Notebooks/tree/master/HACL-2020

Please also see this handy github "cheatsheet": https://education.github.com/git-cheat-sheet-education.pdf

## UNDERSTANDING THE DATA

* All of the data used for this notebook and the **COVIDMINDER** app is located in the `data/` directory hierarchy, mostly under `data/csv/` and `data/csv/time_series/`
* Feel free to browse the directory tree using RStudio or on the web via github
* Most of the nationwide data is under `data/csv`
* Most of the nationwide time series data and New York per-county data is under `data/csv/time_series/`
    * This layout may change
* We will be expanding our data archive, esp. using social determinatns data from <b>County Health Rankings</b>
    * See https://www.countyhealthrankings.org/explore-health-rankings/measures-data-sources/2020-measures 

## WHAT SHOULD I WORK ON?!?!?

* Choose an existing issue: https://github.com/TheRensselaerIDEA/COVID-Notebooks/issues
    * Ask questions about the issue in the "Comments" section of that issue
    * "Claim" the issue by self-assigning 
* OR, create your own issue; click "New Issue"
* OR, ask Prof. Bennett or Dr. Erickson for an issue to work on!

## HOW TO SHARE YOUR WORK OR ASK QUESTIONS

* For class-related questions, use the `data-incite-2020` slack channel under The Rensselaer IDEA
* For COVIDMINDER questions use the `idea_covidminder` stack channel 
    * Contact Dr. Erickson to be added (if you haven't already)

## WHAT TO INCLUDE IN YOUR NOTEBOOK
* The code in your notebook should be well commented
* You should include thorough discussions of your methods and explanations of your results in the **markdown** sections (the non-code sections)
* You should always include enough detail for someone else to reproduce (and re-use!) your work!
* **ALWAYS* include useful titles and legends for your plots and tables!
* Nicely-formatted tables are always a bonus
* Your "default" for knitting notebooks should be HTML; this makes it easier to view and reduces errors. 

<hr>

```{r echo = F, results = 'hide',  warning=FALSE, message=FALSE}
### Text used for captions and explanations

ldi_explanation_text <- "<h4><b>EXPLANATION OF VISUALIZATIONS:</b></h4> The goal of these visualizations is to examine 
nationwide disparities in COVID-19 factors having to do with risks, mediations (e.g. testing, hospital beds), 
and outcomes (e.g. deaths, cases). A common measure, the <i>disparity index</i> is used to represent the 
difference between the observed rate in the state and some baseline rate.</p>
<p>The advantage of the disparity index is that represents how far off a target standard the observed rate is. </p>
<p>Mathematically,
<code>DI = log(x/y)</code> or <code>DI = log(y/x)</code>
depending upon whether being above or below the target is preferred. 
<ul>
<li>In the case of hospital beds or rate of testing, <i>x</i> would be some state's rate , and <i>y</i> would be the US rate or some rate 
we're comparing against (e.g. South Korea's testing or Italy's hospital beds).</li>
<li>In the case of mortality rates, <i>x</i> would be the target rate (e.g. some national rate, including the US), and <i>y</i> would be the individual state's rate.</li>
</ul>
</p>
<p>&nbsp;</p>"

rpi_accessibility_link <- "<div class='center'><p><a href='https://info.rpi.edu/statement-of-accessibility'>Rensselaer Statement of Accessibility</a></p></div>"

### OUTCOMES: COVID-19 Mortality Rates (USA)
us_mortality_rates_text <- "<h4><b>How do COVID-19 mortality rates compare across the United States?</b></h4>
<i>This map compares the COVID-19 mortality rates of individual states with the US rate.
This map is updated daily.</i><br><br>
Here, <span style='color:#b2182b'><b>shades of red</b></span> indicate that a 
state's COVID-19 mortality rate is higher than the US rate<br><br> 
Data source: <a href='https://bit.ly/3dMWRP6'>JHU daily reports</a> (04-07-2020)"

### MEDIATION: COVID-19 Testing (USA)

us_testing_rates_text <- "<h4><b>How do COVID-19 testing rates across the US compare with South Korea?</b></h4>
<i>This map compares rates of COVID-19 tssting in US states vs South Korea's testing rate. 
This map is updated daily.</i><br><br>
Here, <span style='color:#b2182b'><b>shades of red</b></span> indicate that a 
state's testing rate is lower than the South Korean rate<br><br>
Data source: <a href='https://covidtracking.com/api'>The COVID Tracking Project daily reports</a> (04-07-2020)"

### MEDIATION: Hospital Beds (USA)

us_hospital_beds_text <- "<h4><b>How does the availability of hospital beds across the United States compare with Italy?</b></h4>
<i>This map compares the availability of hospital beds in US states vs the rate in Italy (3.2 beds/1000). 
This map uses recent historical figures and does not reflect 'surge' capacity.</i><br><br>
Here, <span style='color:#b2182b'><b>shades of red</b></span> indicate that a 
state's hospital bed availablity is lower than the rate in <b>Italy</b><br/><br>
Data sources: <br/><a href='https://data.oecd.org/healtheqt/hospital-beds.htm'> Organisation for Economic Co-operation and Development</a>
and <a href='https://bit.ly/2V0CYLU'>Kaiser Family Foundation</a>"

### DETERMINANT: Cardiovascular Diseases (USA)

us_cardio_deaths_rates_text <- "<h4><b>How do cardiovascular mortality rates across the US compare with the national average?</b></h4>
<i>The map compares individual state mortality rates related to cardiovascular diseases (per 100k)
with the US rate. In recent literature, COVID-19 risk has been
linked to certain cardiovascular diseases, including hypertension. 
This map uses recent historical figures. </i><br><br>
Here, <span style='color:#b2182b'><b>shades of red</b></span> indicate that a 
state's mortality rate from total cardiovascular diseases is 
<b>higher</b> than the US rate<br/><br>
Data source: <br/><a href='https://bit.ly/2V1Zl3I'>CDC (2017)</a>"

us_diabetes_text <- "<div style='font-weight:bold;line-height:1.3;'>
Determinant: What are the disparities between states in rate of diabetes patients 
per 100k population per state when compared to the average United States rate? </div><br>
<div style='font-size:90%;line-height:1.2;'>
Diabetes puts patients at increased risk of contracting and dying from COVID-19, 
so areas with higher diabetes rates may face increased COVID-19 burdens. <br><br>
The  rate of diabetes deaths per 100k in a state is<br>
<div>&nbsp;&nbsp;&nbsp;<span style='background: #BD0026; border-radius: 50%; font-size: 11px; opacity: 0.7;'>&nbsp&nbsp&nbsp&nbsp</span><strong> Higher</strong> than US avg. rate for disparity index &gt; 0.2</div>
<div>&nbsp;&nbsp;&nbsp;<span style='background: #ffffff; border-radius: 50%; font-size: 11px; opacity: 0.7;'>&nbsp&nbsp&nbsp&nbsp</span><strong> About equal</strong> to US avg. rate for -0.2 &lt;disparity index &lt; 0.2</div>
<div>&nbsp;&nbsp;&nbsp;<span style='background: #253494; border-radius: 50%; font-size: 11px; opacity: 0.7;'>&nbsp&nbsp&nbsp&nbsp</span><strong> Lower</strong> than US avg. rate for disparity index &lt; -0.2</div>
<i>Darker shades indicate greater disparity.</i><br><br>
<strong>Diabetes Rate</strong> = number of diabetic patients per 100K population <br>
<strong>Diabetes Disparity Index</strong> = log(Diabetes Rate in state/average Diabetes Rate in US)<br>
<strong>Date: </strong> 2020<br><br>
<b>DATA SOURCE:</b> <a href='https://bit.ly/34mYLBP'>County Health Rankings</a> and 
<a href='https://bit.ly/2V1Zl3I'>CDC</a><br>
</div>"
                            
NY_diabetes_text <- "<div style='font-weight:bold;line-height:1.3;'>
Determinant: What are the disparities between New York counties in the rate 
of diabetes patients per 100k population when compared to the average United 
States rate?</div><br>
<div style='font-size:90%;line-height:1.2;'>
Diabetes puts patients at increased risk of contracting and dying from COVID-19, 
so areas with higher diabetes rates may face increased COVID-19 burdens. <br><br>
The  rate of diabetes patients per 100k in a county  is<br>
<div>&nbsp;&nbsp;&nbsp;<span style='background: #BD0026; border-radius: 50%; font-size: 11px; opacity: 0.7;'>&nbsp&nbsp&nbsp&nbsp</span><strong> Higher</strong> than US avg. rate for disparity index &gt; 0.2</div>
<div>&nbsp;&nbsp;&nbsp;<span style='background: #ffffff; border-radius: 50%; font-size: 11px; opacity: 0.7;'>&nbsp&nbsp&nbsp&nbsp</span><strong> About equal</strong> to US avg. rate for -0.2 &lt; disparity index &lt; 0.2</div>
<div>&nbsp;&nbsp;&nbsp;<span style='background: #253494; border-radius: 50%; font-size: 11px; opacity: 0.7;'>&nbsp&nbsp&nbsp&nbsp</span><strong> Lower</strong> than US avg. rate for disparity index &lt; -0.2</div>
<i>Darker shades indicate greater disparity.</i><br><br>
<strong>Diabetes Rate</strong> = number of diabetic patients  per 100K population <br>
<strong>Diabetes Disparity Index</strong> = log(Diabetes Rate in state/average Diabetes Rate in US)<br>
<strong>Date: </strong> 2020<br><br>
<b>DATA SOURCE:</b> <a href='https://bit.ly/34mYLBP'>County Health Rankings</a> and 
<a href='https://bit.ly/2V1Zl3I'>CDC</a><br>
</div>"

### OUTCOME: COVID-19 mortality rates (NY)

NY_mortality_rates_text <- "<h4><b>How do COVID-19 mortality rates compare across New York State?</b></h4>
<i>This map compares the COVID-19 mortality rates of NY counties with the NY average. 
This map is updated daily. </i><br><br>
Here, <span style='color:#b2182b'><b>shades of red</b></span> indicate that a 
county's COVID-19 mortality rate is higher than the NY rate.<br>
Data source: <a href='https://bit.ly/3dMWRP6'>JHU daily reports</a> (04-06-2020)"

### OUTCOME: COVID-19 Cases (NY)

NY_covid_cases_text <- "<h4><b>How do COVID-19 cases compare across New York State?</b></h4>
<i>This map compares the COVID-19 cases for NY counties with the NY average. 
This map is updated daily. </i><br><br>
Here, <span style='color:#b2182b'><b>shades of red</b></span> indicate that a 
county's COVID-19 case count is higher than the NY rate.<br>
Data source: <a href='https://bit.ly/3dMWRP6'>JHU daily reports</a> (04-06-2020)"

### OUTCOME: COVID-19 Cases Time Series (NY)

NY_covid_cases_TS_text <- "<h4><b>How have COVID-19 cases increased across New York State?</b></h4>
<i>This plot shows the growth of COVID-19 cases across NY counties since early 2020. 
This data is updated daily. </i><br><br>
Mouse over the plot to identify individual counties.<br>
Data source: <a href='https://bit.ly/3dMWRP6'>JHU daily reports</a> (04-06-2020)"

```

`r ldi_explanation_text`

`r us_mortality_rates_text`

