---
title: "COVIDMINDER Insurance Data Notebook"
author: "Nicole Schwartzbard"
date: "4/20/2020"
output:
  html_document: default
  pdf_document: default
---
<h5><b>Reasoning of Looking at Insurance:</b></h5>
I wanted to study insurance rates since more people then normally expected now need care, and therefore insurance. For the US government that now becomes very expensive and for those uninsured it is now dangerous.This data is from the uninsured_by_state csv file but I don't know where that is from. There is duplicate work, formating the data, in each code chunk to keep the chunks separate if not all will be used.

<h5>Methods:</h5>
I used the same code from the COVIDMINDER Bootcamp Notebook so that the maps would look the same, though I did edit the borders because I didn't like the white dashed outlines. I used the uninsured_by_state csv to first display the actual percentages, and then I calculated the disparity index and displayed those.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, results = 'hide',  warning=FALSE, message=FALSE}
#### Library and Data Imports ####
r = getOption("repos")
r["CRAN"] = "http://cran.rstudio.com"
options(repos = r)

## Load essential R packages
source("modules/Source.R")

## Load stored data (see `data` subdirectory)
source("modules/data_load.R")

## Create dataframes; perform desparity index calcuations; prep for plotting
source("modules/preprocessing.R")

```


```{r echo = F, results = 'hide',  warning=FALSE, message=FALSE}
### Text used for captions and explanations

Insurance_explanation_text <- "<h4><b>EXPLANATION OF VISUALIZATIONS:</b></h4> The goal of these visualizations is to examine nationwide differences in how people across the US are insured. The graphs that use Disparity Index compare state rates to the united states rate, where disparity is calculated as log(x/y), where x is the state rate and y is the United States rate.
<i>Those insured by Medicare and Medicaid will be limited by the money the US government has, while those uninsured are at risk of being unable to pay for care.</i>"


### US Uninsured Percentage

US_uninsured_text = "<h4><b>How does the perecentage of people uninsured vary across the United States?</b></h4>
<i> This plot shows the differences of percentage of people uninsured across the US. This data is from _______.</i>"
#I don't know where this data was from, uses the uninsured_by_state data

### US Medicaid Percentage

US_medicaid_text = "<h4><b>How does the perecentage of people insured by Medicaid vary across the United States?</b></h4><i> This plot shows the differences of percentage of people uninsured across the US. This data is from _______.</i>"
#I don't know where this data was from, uses the uninsured_by_state data

### US Medicare Percentage

US_medicare_text = "<h4><b>How does the perecentage of people insured by Medicare vary across the United States?</b></h4><i> This plot shows the differences of percentage of people uninsured across the US. This data is from _______.</i>"
#I don't know where this data was from, uses the uninsured_by_state data


###US Uninsured DI
US_uninsured_DI_text = "<h4><b>How does the perecentage of people uninsured compare across the United States?</b></h4>
<i> This plot shows the disparity index of percentage people uninsured across the US. Here red indicates that there is a higher percentage than average who are uninsured. The disparity index is calculated by ln(x/y), where x is the state data and y is the country data. This data is from _______.</i>"
#I don't know where this data was from, uses the uninsured_by_state data


###US Medicaid DI
US_medicaid_DI_text = "<h4><b>How does the perecentage of people insured by medicaid compare across the United States?</b></h4>
<i> This plot shows the disparity index of percentage people insured by medicaid across the US. Here red indicates that there is a higher percentage than average who are insured by medicaid. The disparity index is calculated by ln(x/y), where x is the state data and y is the country data. This data is from _______.</i>"
#I don't know where this data was from, uses the uninsured_by_state data


###US Medicare DI
US_medicare_DI_text = "<h4><b>How does the perecentage of people insured by medicare compare across the United States?</b></h4>
<i> This plot shows the disparity index of percentage people insured by medicare across the US. Here red indicates that there is a higher percentage than average who are insured by medicare. The disparity index is calculated by ln(x/y), where x is the state data and y is the country data.This data is from _______.</i>"
#I don't know where this data was from, uses the uninsured_by_state data

### NY Medicaid DI
NY_medicaid_DI_text = "<h4><b> How does the perecentage of people insured by medicaid compare across New York</h4></b><i> This plot shows the disparity index of percentage people insured by medicaid across NY. Here red indicates that there is a higher percentage than average who are insured by medicaid. The disparity index is calculated by ln(x/y), where x is the county data and y is the state data. This data is from <a href='https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/2020/docs/en03_20.pdf'>health.ny.gov</a>."

```

`r Insurance_explanation_text`

`r US_uninsured_text`

```{r echo = F}
#Get Uninsured data and format column names correctly
Uninsured = read.csv('data/csv/uninsured_by_state.csv', row.names = NULL)
colnames(Uninsured) = c("NAME", "Status", "Employer", "Non-group", "Medicaid", "Medicare", "Military", "Uninsured", "number_uninsured")

#match order with states in state data so that the state name is accurate on the map
Uninsured = Uninsured[match(states$NAME, Uninsured$NAME),]


 colors <- c("grey","#426C85","#67a9cf","#d1e5f0","#f7f7f7","#fddbc7","#ef8a62","#b2182b")
    bins <- c(0, 0.025, 0.05, 0.075, 0.1, 0.125, 0.15, 0.175, 0.2)
    pal2 <- leaflet::colorBin(colors, domain = Uninsured$Uninsured, bins = bins, reverse=FALSE)
    labels2 <- sprintf(
      "<strong>%s</strong><br/>
      Percent of People Uninsured: %.9g",
      Uninsured$NAME, Uninsured$Uninsured
    ) %>% lapply(htmltools::HTML)
    
    leaflet(states.shapes) %>%
      setView(-96, 37.8, 4) %>% 
      addPolygons(
        fillColor = ~pal2(Uninsured$Uninsured),
        weight = 2,
        opacity = 1,
        color = "grey",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels2,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~Uninsured$Uninsured, 
                opacity = 0.7, 
                title = "US Uninsured Percentage",
                position = "bottomright"
                ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light",
        accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
```


`r US_medicaid_text`

```{r echo = F}
#Get Uninsured data and format column names correctly
Uninsured = read.csv('data/csv/uninsured_by_state.csv', row.names = NULL)
colnames(Uninsured) = c("NAME", "Status", "Employer", "Non-group", "Medicaid", "Medicare", "Military", "Uninsured", "number_uninsured")

#match order with states in state data so that the state name is accurate on the map
Uninsured = Uninsured[match(states$NAME, Uninsured$NAME),]


#match order with states in state data
Uninsured = Uninsured[match(states$NAME, Uninsured$NAME),]


 colors <- c("grey","#426C85","#67a9cf","#d1e5f0","#f7f7f7","#fddbc7","#ef8a62","#b2182b")
    bins <- c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30, 0.35)
    pal2 <- leaflet::colorBin(colors, domain = Uninsured$Uninsured, bins = bins, reverse=FALSE)
    labels2 <- sprintf(
      "<strong>%s</strong><br/>
      Percentage of People insured by Medicaid: %.9g",
      Uninsured$NAME, Uninsured$Medicaid
    ) %>% lapply(htmltools::HTML)
    
    leaflet(states.shapes) %>%
      setView(-96, 37.8, 4) %>% 
      addPolygons(
        fillColor = ~pal2(Uninsured$Medicaid),
        weight = 2,
        opacity = 1,
        color = "grey",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels2,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~Uninsured$Medicaid, 
                opacity = 0.7, 
                title = "US Insured By Medicaid",
                position = "bottomright"
                ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light",
        accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
```


`r US_medicare_text`

```{r echo = F}
#Get Uninsured data and format column names correctly
Uninsured = read.csv('data/csv/uninsured_by_state.csv', row.names = NULL)
colnames(Uninsured) = c("NAME", "Status", "Employer", "Non-group", "Medicaid", "Medicare", "Military", "Uninsured", "number_uninsured")

#match order with states in state data so that the state name is accurate on the map
Uninsured = Uninsured[match(states$NAME, Uninsured$NAME),]


 colors <- c("grey","#426C85","#67a9cf","#d1e5f0","#f7f7f7","#fddbc7","#ef8a62","#b2182b")
    bins <- c(0, 0.075, 0.10, 0.125, 0.15, 0.175, 0.2)
    pal2 <- leaflet::colorBin(colors, domain = Uninsured$Uninsured, bins = bins, reverse=FALSE)
    labels2 <- sprintf(
      "<strong>%s</strong><br/>
      Percentage of People insured by Medicare: %.9g",
      Uninsured$NAME, Uninsured$Medicare
    ) %>% lapply(htmltools::HTML)
    
    leaflet(states.shapes) %>%
      setView(-96, 37.8, 4) %>% 
      addPolygons(
        fillColor = ~pal2(Uninsured$Medicare),
        weight = 2,
        opacity = 1,
        color = "grey",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels2,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~Uninsured$Medicare, 
                opacity = 0.7, 
                title = "US Insured By Medicare",
                position = "bottomright"
                ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light",
        accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
```


`r US_uninsured_DI_text`

```{r echo = F}
#Get Uninsured data and format column names correctly
Uninsured = read.csv('data/csv/uninsured_by_state.csv', row.names = NULL)
colnames(Uninsured) = c("NAME", "Status", "Employer", "Non-group", "Medicaid", "Medicare", "Military", "Uninsured", "number_uninsured")

#Get America's Average Uninsured
Avg = Uninsured[Uninsured$NAME == "United States",]$Uninsured

#match order with states in state data so that the state name is accurate on the map
Uninsured = Uninsured[match(states$NAME, Uninsured$NAME),]

#Calculate Disparity Index
Uninsured$DI = log(Uninsured$Uninsured/Avg)

colors <- c("grey","#426C85","#67a9cf","#d1e5f0","#f7f7f7","#fddbc7","#ef8a62","#b2182b")
    bins <- c(-1.2, -0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8)
    #bins = c(-5, -3, -2, -1, -0.2, 0.2, 1, 2, 3, 5)
    pal2 <- leaflet::colorBin(colors, domain = Uninsured$DI, bins = bins, reverse=FALSE)
    labels2 <- sprintf(
      "<strong>%s</strong><br/>
      People Uninsured DI: %.9g",
      Uninsured$NAME, Uninsured$DI
    ) %>% lapply(htmltools::HTML)
    
    leaflet(states.shapes) %>%
      setView(-96, 37.8, 4) %>% 
      addPolygons(
        fillColor = ~pal2(Uninsured$DI),
        weight = 2,
        opacity = 1,
        color = "grey",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels2,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~Uninsured$DI, 
                opacity = 0.7, 
                title = "US Uninsured Disparity Index",
                position = "bottomright"
                ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light",
        accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))

```

`r US_medicaid_DI_text`

```{r echo = F}
#Get Uninsured data and format column names correctly
Uninsured = read.csv('data/csv/uninsured_by_state.csv', row.names = NULL)
colnames(Uninsured) = c("NAME", "Status", "Employer", "Non-group", "Medicaid", "Medicare", "Military", "Uninsured", "number_uninsured")

#Get America's Average Medicaid
Avg = Uninsured[Uninsured$NAME == "United States",]$Medicaid

#match order with states in state data so that the state name is accurate on the map
Uninsured = Uninsured[match(states$NAME, Uninsured$NAME),]

#Calculate Disparity Index
Uninsured$DI = log(Uninsured$Medicaid/Avg)

colors <- c("grey","#426C85","#67a9cf","#d1e5f0","#f7f7f7","#fddbc7","#ef8a62","#b2182b")
    bins <- c(-0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6)
    #bins = c(-5, -3, -2, -1, -0.2, 0.2, 1, 2, 3, 5)
    pal2 <- leaflet::colorBin(colors, domain = Uninsured$DI, bins = bins, reverse=FALSE)
    labels2 <- sprintf(
      "<strong>%s</strong><br/>
      People Medicaid DI: %.9g",
      Uninsured$NAME, Uninsured$DI
    ) %>% lapply(htmltools::HTML)
    
    leaflet(states.shapes) %>%
      setView(-96, 37.8, 4) %>% 
      addPolygons(
        fillColor = ~pal2(Uninsured$DI),
        weight = 2,
        opacity = 1,
        color = "grey",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels2,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~Uninsured$DI, 
                opacity = 0.7, 
                title = "US Medicaid Disparity Index",
                position = "bottomright"
                ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light",
        accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))

```


`r US_medicare_DI_text`

```{r echo = F}
#Get Uninsured data and format column names correctly
Uninsured = read.csv('data/csv/uninsured_by_state.csv', row.names = NULL)
colnames(Uninsured) = c("NAME", "Status", "Employer", "Non-group", "Medicaid", "Medicare", "Military", "Uninsured", "number_uninsured")

#Get America's Average Medicare
Avg = Uninsured[Uninsured$NAME == "United States",]$Medicare

#match order with states in state data so that the state name is accurate on the map
Uninsured = Uninsured[match(states$NAME, Uninsured$NAME),]

#Calculate Disparity Index
Uninsured$DI = log(Uninsured$Medicare/Avg)

colors <- c("grey","#426C85","#67a9cf","#d1e5f0","#f7f7f7","#fddbc7","#ef8a62","#b2182b")
    bins <- c(-0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6)
    #bins = c(-5, -3, -2, -1, -0.2, 0.2, 1, 2, 3, 5)
    pal2 <- leaflet::colorBin(colors, domain = Uninsured$DI, bins = bins, reverse=FALSE)
    labels2 <- sprintf(
      "<strong>%s</strong><br/>
      Medicare DI: %.9g",
      Uninsured$NAME, Uninsured$DI
    ) %>% lapply(htmltools::HTML)
    
    leaflet(states.shapes) %>%
      setView(-96, 37.8, 4) %>% 
      addPolygons(
        fillColor = ~pal2(Uninsured$DI),
        weight = 2,
        opacity = 1,
        color = "grey",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels2,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~Uninsured$DI, 
                opacity = 0.7, 
                title = "US Medicare Disparity Index",
                position = "bottomright"
                ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light",
        accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))

```


`r NY_medicaid_DI_text`

```{r echo = F}
#read in data
medicaid_users = read.csv('data/csv/ny_medicaid_03_20.csv')

#head(medicaid_users)

#get just totals
medicaid_users = medicaid_users[medicaid_users$Plan.Name == "TOTALS:",]

#match data to NY.data order
medicaid_users = medicaid_users[match(NY.data$County, medicaid_users$County), ]

medicaid_users$Xnum = as.numeric(paste(medicaid_users$X))
medicaid_users$County = as.character(paste(medicaid_users$County))
#Calculate percentages of medicaid users

medicaid_users$percent = medicaid_users$Xnum/NY.data$Population

#Calculate Disparity Index
Avg = Uninsured[Uninsured$NAME == "New York",]$Medicaid

NY.data$DI = log(medicaid_users$percent/Avg)


    colors <- c("grey","#426C85","#67a9cf","#d1e5f0","#f7f7f7","#fddbc7","#ef8a62","#b2182b")
    bins <- c(-1.2, -1, -0.8, -.6, -.4, -.2, 0, .2, .4, .6)
    pal2 <- leaflet::colorBin(colors, domain = NY.data$death_rate_ldi , bins = bins, reverse=FALSE)
    
    #domain = NY.data$death_rate_ldi
    NY.shape$county_fips <- paste(as.data.frame(NY.shape)$STATEFP, as.data.frame(NY.shape)$COUNTYFP, sep = '')
    NY.data <- dplyr::left_join(as.data.frame(NY.shape), as.data.frame(NY.data), by = c("county_fips" = "FIPS"))
    
    labels <- sprintf(
      "<strong>%s</strong><br/>
      Insured by Medicaid Disparity Index: %.9g",
      NY.data$County, NY.data$DI
    ) %>% lapply(htmltools::HTML)

    leaflet(NY.shape) %>%
      setView(-76.071782, 42.991989, 6) %>%  # Set to the geographic center of NY
      addPolygons(
        fillColor = ~pal2(NY.data$DI ),
        weight = 2,
        opacity = 1,
        color = "grey",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~NY.data$death_rate_ldi, 
                opacity = 0.7, 
                title = "Disparity Index<br/>NY Medicaid",
                position = "bottomright"
      ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light",
        accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
```