---
title: "COVIDMINDER Obesity Data Notebook"
author: "Nicole Schwartzbard"
date: "4/21/2020"
output:
  html_document: default
  pdf_document: default
---
<h5><b>Reasoning of Looking at Obesity:</b></h5>
This can be useful determinats for Covid-19. I could only find data based off of students for actual obesity, but I found data for limited access to healthy foods which is a contributing factor for obesity so I used that.


<h5>Methods:</h5>
I used the same code from the COVIDMINDER Bootcamp Notebook so that the maps would look the same, though I did edit the borders because I didn't like the white dashed outlines. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, results = 'hide',  warning=FALSE, message=FALSE}
#### Library and Data Imports ####
r = getOption("repos")
r["CRAN"] = "http://cran.rstudio.com"
options(repos = r)

## Load essential R packages
source("modules/Source.R")

## Load stored data (see `data` subdirectory)
source("modules/data_load.R")

## Create dataframes; perform desparity index calcuations; prep for plotting
source("modules/preprocessing.R")

```


```{r echo = F, results = 'hide',  warning=FALSE, message=FALSE}
### Text used for captions and explanations

### NY Asthma DI
NY_food_DI_text = "<h4><b> How does the percentage of people who are healthy food insecure compare across New York</h4></b><i> This plot shows the disparity index of percentage of people who are healthy food insecure across NY. Here red indicates that there is a higher percentage than average who don't have access. The disparity index is calculated by ln(x/y), where x is the county data and y is the state data. This data is from <a href='https://www.countyhealthrankings.org/app/new-york/2020/downloads'>www.countyhealthrankings.org</a>."

```


`r NY_food_DI_text`

```{r echo = F}
#read in data
food = read.csv('county_data.csv')
food$percent = as.numeric(paste(food$X..Limited.Access.to.Healthy.Foods))

Avg = food[food$County == "",]$percent

food = food[match(NY.data$County, food$County), ]


NY.data$DIf = log(food$percent/Avg)

  colors <- c("grey","#426C85","#67a9cf","#d1e5f0","#f7f7f7","#fddbc7","#ef8a62","#b2182b")
    bins <- c( -Inf, -2, -1.5, -.5, -.2, 0, .2, .5, 1, 1.5, 2.0)
    pal2 <- leaflet::colorBin(colors, domain = NY.data$death_rate_ldi , bins = bins, reverse=FALSE)
    
    #domain = NY.data$death_rate_ldi
    NY.shape$county_fips <- paste(as.data.frame(NY.shape)$STATEFP, as.data.frame(NY.shape)$COUNTYFP, sep = '')
    NY.data <- dplyr::left_join(as.data.frame(NY.shape), as.data.frame(NY.data), by = c("county_fips" = "FIPS"))
    
    labels <- sprintf(
      "<strong>%s</strong><br/>
      Limited Healthy Food Access Disparity Index: %.9g",
      NY.data$County, NY.data$DIf
    ) %>% lapply(htmltools::HTML)

    leaflet(NY.shape) %>%
      setView(-76.071782, 42.991989, 6) %>%  # Set to the geographic center of NY
      addPolygons(
        fillColor = ~pal2(NY.data$DIf ),
        weight = 2,
        opacity = 1,
        color = "grey",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~NY.data$death_rate_ldi, 
                opacity = 0.7, 
                title = "Disparity Index<br/>NY Limited Healthy Food Access",
                position = "bottomright"
      ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light",
        accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
```


