---
title: "COVIDMINDER Incident and Mortality Notebook"
author: "Michél McFadden"
date: "5/4/2020"
output:
  html_document: default
  pdf_document: default
---
<h5><b>Why Incident and Mortality Is Important :</b></h5>
Looking at Incident and Mortality is important because it can help predict the future of the state. To flatten the curve states want to keep their incident rates low over time. It may be possible to predict when states might open by looking at their incident rates and mortality rates. It's reasonable to assume that states that have low rates for both will open back up sooner than states with higher rates. 

<h5>Methods:</h5>
INSERT DESCRIPTION


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(gplots)
library(ggplot2)
library(ggbiplot)
library(rlang)
```

```{r echo = F, results = 'hide',  warning=FALSE, message=FALSE}
#### Library and Data Imports ####
r = getOption("repos")
r["CRAN"] = "http://cran.rstudio.com"
options(repos = r)

## Load essential R packages
source("modules/Source.R")

## Load stored data (see `data` subdirectory)
source("modules/data_load.R")

## Create dataframes; perform desparity index calcuations; prep for plotting
source("modules/preprocessing.R")
```


```{r echo = F, results = 'hide',  warning=FALSE, message=FALSE}
### Text used for captions and explanations

Graph_explanation_text <- "<h4><b>ABOUT THE VISUALIZATIONS:</b></h4> These maps go to show the different rates for states. The Mortality Rate graph uses the data untouched directly from the CSV and the Incident Graph takes the log(X) of each states incident rate (the amount of people that catch coronvirus divided by the population value."


# Mortality Graph Text
mortality_text = "<h4><b>What does the mortality rate look like across the United States?</b></h4>
<i> This plot shows the different mortality rates across the US. This data is from JHU's Github repo.</i>"

# Incident Rate Text 
incident_text <- "<h4><b> What does the incident rate look like across the United States? </b></h4>
<i> This plot shows the different incident rates across the US. 'Incident rate' is defined as the amount of people who catch coronavirus in a given state divided by the population of that state. This data is from JHU's Github repo. </i>"

```

`r Graph_explanation_text`

`r mortality_text`

```{r echo = F}
#Get Uninsured data and format column names correctly
COVID.df = read.csv('data/csv/05-01-2020.csv', row.names = NULL)

# Removeing the unnecessary features and places
COVID.df$Country_Region <- NULL
COVID.df$Last_Update <- NULL
COVID.df$Lat <- NULL
COVID.df$Long_ <- NULL
COVID.df$FIPS <- NULL
COVID.df$UID <- NULL
COVID.df$ISO3 <- NULL
COVID.df$Confirmed <- NULL
COVID.df$Deaths <- NULL
COVID.df$Recovered <- NULL
COVID.df$Active <- NULL
COVID.df$People_Hospitalized <- NULL
COVID.df$People_Tested <- NULL
COVID.df <- COVID.df[-c(10, 14, 40, 53), ]
colnames(COVID.df) <- c("NAME", "Incident", "Mortality", "Testing", "Hospitalization")


# Setting NAs to 0
COVID.df[is.na(COVID.df)] <- 0
COVID.df$Incident <- log(COVID.df$Incident + 1)

#match order with states in state data so that the state name is accurate on the map
COVID.df = COVID.df[match(states$NAME, COVID.df$NAME),]


 colors <- c("#ffff00","#ff7f50","#d1e5f0","#fddbc7","#ff8100","#ff8d00","#b2182b")
    bins <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
    pal2 <- leaflet::colorBin(colors, domain = COVID.df$Mortality, bins = bins, reverse=FALSE)
    labels2 <- sprintf(
      "<strong>%s</strong><br/>
      Mortality Rate: %.9g",
      COVID.df$NAME, COVID.df$Mortality
    ) %>% lapply(htmltools::HTML)
    
    leaflet(states.shapes) %>%
      setView(-96, 37.8, 4) %>% 
      addPolygons(
        fillColor = ~pal2(COVID.df$Mortality),
        weight = 2,
        opacity = 1,
        color = "grey",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels2,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~COVID.df$Mortality, 
                opacity = 0.7, 
                title = "US Mortality Rates by State",
                position = "bottomright"
                ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light",
        accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
```


`r incident_text`

```{r echo = F}

 colors <- c("#fffcb3","#fff96a", "#ffff00","#fddbc7","#ff8100","#ff8d00","#b2182b")
    bins <- c(0, 1, 2, 3, 4, 5, 6, 7, 8)
    pal2 <- leaflet::colorBin(colors, domain = COVID.df$Incident, bins = bins, reverse=FALSE)
    labels2 <- sprintf(
      "<strong>%s</strong><br/>
      Incident Rate: %.9g",
      COVID.df$NAME, COVID.df$Incident
    ) %>% lapply(htmltools::HTML)
    
    leaflet(states.shapes) %>%
      setView(-96, 37.8, 4) %>% 
      addPolygons(
        fillColor = ~pal2(COVID.df$Incident),
        weight = 2,
        opacity = 1,
        color = "grey",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels2,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~COVID.df$Incident, 
                opacity = 0.7, 
                title = "US Incident Rates by State",
                position = "bottomright"
                ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light",
        accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
```

<i> New York had such inflated values that including them in the analysis would lead to overfitting the data. New York data was removed from the analysis techniques below </i>

```{r echo = F}
# Downloading the COVID data into ta data frame
COVID.df <- read.csv("data/csv/05-01-2020.csv", row.names = 1)


# Removeing the unnecessary features and places
COVID.df$Country_Region <- NULL
COVID.df$Last_Update <- NULL
COVID.df$Lat <- NULL
COVID.df$Long_ <- NULL
COVID.df$FIPS <- NULL
COVID.df$UID <- NULL
COVID.df$ISO3 <- NULL
COVID.df$Confirmed <- NULL
COVID.df$Deaths <- NULL
COVID.df$Recovered <- NULL
COVID.df$Active <- NULL
COVID.df$People_Hospitalized <- NULL
COVID.df$People_Tested <- NULL
COVID.df <- COVID.df[-c(10, 14, 37, 40, 53), ]
colnames(COVID.df) <- c("Incident", "Mortality", "Testing", "Hospitalization")

# Setting NAs to 0
COVID.df[is.na(COVID.df)] <- 0

# Converting the data frame and scaling the resulting matrix
COVID.matrix <- data.matrix(COVID.df)
COVID.matrix <- scale(COVID.matrix, center = TRUE, scale = TRUE)

# Principle Component Analysis and Screeplot 
COVID.pca <- prcomp(COVID.matrix, retx=TRUE, center=TRUE, scale=TRUE)
screeplot(COVID.pca, type = "lines", main = 'Explanation of Variance for State COVID')
```


# A look at the heatmap values for the different cluster of states. 

* Cluster A is characterized by it's low testing, incident, and mortality rates and medium hospitalization rate.

* Cluster B is characterized by it's low hospitalization rate, medium testing and mortality rates, and high incident rate.

* Cluster C is characterized by it's low rates for all 4 features.

```{r echo = F}
# K Means Clustering
set.seed(300)
COVID.km <-kmeans(COVID.matrix, 3)
COVID.km$cluster <- as.factor(COVID.km$cluster)

Clusters <- COVID.km$cluster
levels(Clusters) <- c("A", "B", "C")

my_palette <- colorRampPalette(c("yellow", "goldenrod2", "coral1"))(n = 500)

heatmap.2(COVID.km$centers, col=my_palette, scale=NULL)
```


# What information can we get from looking at a biplot?
<h4><b> A biplot shows the characteristics of a cluster in relation to one another </b></h4>

* For example we see Cluster A's medium hospitalization rate characterized by the blue states plotted higher on the Hospitalization Rate eigenvector

* We can see Cluster B's high incident rate by the golden states with large placements on the Incident Rate eigenvector

* Cluster C's low values across all rates are demonstrated by it's clustering close to the origin of the biplot

```{r echo = F}
ggbiplot(COVID.pca,choices=c(1,2),
                labels=rownames(COVID.matrix), #show point labels
                var.axes=TRUE, # Display axes 
                ellipse = FALSE, # Don't display ellipse
                obs.scale=.01,
                groups = Clusters) + scale_color_manual(values=c("dodgerblue1", "goldenrod2", "coral1")) + labs(title = "Biplot for COVID Data Features")
```


