---
title: "State Report Cards"
author: "Jose Figueroa"
date: 06/02/20
output: html_notebook
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
knitr::opts_knit$set(root.dir = "../")
```

```{r imports, echo = F, results = 'hide',  warning=FALSE, message=FALSE}
## Load essential R packages
source("modules/Source.R")
source("modules/data_load.R")
source("modules/preprocessing.R")
source("modules/ggplot_gen.R")
source("modules/leaflet_gen.R")
source("modules/gt_gen.R")
```


# Data Exploration

There are two main types of data we work with on COVID**MINDER**. GGplots work with time series data, and Leaflet Plots work with static data points for each state. 

## ggplot datasets

```{r}
covid_TS_counties_long.cases %>% str() %>% print() # County Level data
covid_TS_state_long.cases %>% str() %>% print() # State Level data
covid_TS_US_long.cases %>% str() %>% print() # National Data
```

Metadata:
* countyFIPS - Unique integer ID identifying a United States County. The first 2 numbers can be processed into StateFips
* County - Name of the county. Not Unique.
* State - Two letter state initial. Uniquely Identifies state. 
* stateFIPS -  Unique integer ID identifying a United States state.
* date - Date in which row data was reported. Please note unix date_time format.
* cases - Overall COVID-19 cases as of reported date.
* deaths - Overall COVID-19 deaths as of reported date.
* p_cases - Overall COVID-19 cases per 100k as of reported date.
* p_deaths - Overall COVID-19 deaths per 100k as of reported date.
* diff - Change in COVID-19 cases from previous day. Generally >= 0, though some data report discrepancies may generate negatives.
* p_diff - Change in COVID-19 cases per 100k from previous day. Generally >= 0, though some data report discrepancies may generate negatives.
* d_diff - Change in COVID-19 deaths from previous day. Generally >= 0, though some data report discrepancies may generate negatives.
* p.d_diff - Change in COVID-19 deaths per 100k from previous day. Generally >= 0, though some data report discrepancies may generate negatives.

In `covid_TS_counties_long.cases`, a row is uniquely identified by its `countyFIPS` and `date` columns.
In `covid_TS_state_long.cases`, a row is uniquely identified by its `State` and `date` columns.
In `covid_TS_US_long.cases`, a row is uniquely identified by its `date` column.

Our GGplot functions access these dataframes directly, to report data at the state or national level. They can be used to plot any non-identifying column. Command+Click on the functions to view the code.

```{r}
ggplot.state("NY","p_diff", counties = c("Rensselaer"))
```

```{r}
ggbar.overall("NJ", "p_cases")
```

```{r}
ggplot.US("p_diff", selected.states = c("NY", "CA", "NJ", "GA"), moving.avg.window = 14)
```

```{r}
ggbar.US("p_deaths")
```

## leaflet datasets

```{r}
todays.case.data %>% str() %>% print()
states %>% str() %>% print()
```

Any datapoint with a log disparity index (LDI) can be generated into a leaflet plot. `geo.plot` is currently built to visualize variables that have the following columns:
* `feature name`_rate
* `feature name`_rate_ldi


***Note: Legend depends on CSS stylings to have flat visualization.***
```{r}
geo.plot("NY", "Mortality")
```

```{r}
geo.plot("US", "Testing", reverse = T)
```

# Tab panel example
Here is a small application with a singular tabpanel in a navbar setup. To add to covidminder, simply add a tab panel where it is desired. 

```{r}
ui <- tagList(
    tags$head(includeCSS("www/style.css")),
    navbarPage(
    title = tags$div(class = "title-text", "COVIDMINDER"),
    id = "tab",
    tabPanel( # Separate Tab Panel
      title = "Example Panel 1",
      value = "example", # Should be unique and appended to tabs list in the server code
      fluidRow(tags$h1("Daily New York Cases/100k")),
      fluidRow(plotOutput("ny_output"))
    ),
    navbarMenu( # List of Tab Panels in drop down menu
      title = "Example Menu 1",
               tabPanel(
                 title = "Example Panel 2",
                 value = "example2",
                 fluidRow(tags$h1("Nationwide Testing")),
                 fluidRow(leafletOutput("US_output"))
               ))
    #,
    #tabPanel(...),
    #navbarMenu(tabPanel(...), tabPanel(...)),
    #tabPanel(...)
  )
)
```

```{r}
server <- function(input, output) {
  output$ny_output <- renderPlot({
    ggplot.state(selected.state = "NY", 
                 y.value = "p_diff", 
                 moving.avg.window = 14, 
                 counties = c("Rensselaer", "New York City", "Albany", "Nassau", "Erie"),
                 remove.title = T)
  })
  
  output$US_output <- renderLeaflet({
    geo.plot(state.choice = "US", 
             feature = "Testing",
             reverse = T)
  })
}
```

```{r}
shinyApp(ui = ui, server = server)
```


