---
title: "State Report Cards"
author: "Jose Figueroa"
date: 06/02/20
output: html_notebook
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
knitr::opts_knit$set(root.dir = "../")
```

```{r imports, echo = F, results = 'hide',  warning=FALSE, message=FALSE}
#### Library and Data Imports ####

## Load essential R packages
source("modules/Source.R")

## Load stored data (see `data` subdirectory)
source("modules/data_load.R")

## Create dataframes; perform desparity index calcuations; prep for plotting
source("modules/preprocessing.R")

library(gt)

```

```{r map colors}
colors <- c("#253494","#4575B4", "#74ADD1","#ABD9E9","#f7f7f7","#FDAE61","#F46D43", "#D73027", "#BD0026")
bins <- c(5, 3, 2, 1, .2, -.2, -1, -2, -3, -5)
```

Functions to be generalized
```{r get_ldi}
get_ldi <- function(feature) {
  rate <- paste0(feature,"_rate")
  ldi <- paste0(rate, "_ldi")
  return(c(ldi, rate))
}
```



```{r geo.plot}
geo.plot <- function(state.choice, feature, reverse=F) {
  # Feature: Case, Mortality...
  ldi_feature <- get_ldi(feature)
  
  dataset <- todays.case.data %>%
    filter(State == state.choice)
  shapes <- readRDS(paste("data/shape_files/", state.choice, ".Rds", sep = ""))
  shapes$countyFIPS <- as.numeric(paste(as.data.frame(shapes)$STATEFP, as.data.frame(shapes)$COUNTYFP, sep = ''))
  
  dataset <- dplyr::left_join(as.data.frame(shapes), as.data.frame(dataset), by = c("countyFIPS" = "countyFIPS"))
  
  pal2 <- leaflet::colorBin(colors, domain = dataset[,ldi_feature[1]], bins = bins, reverse=reverse)
    
  labels <- sprintf(
    paste0("<strong>%s</strong><br/>
    COVID-19 ",feature," Rate DI: %.2g<br>
    COVID-19 ",feature," Rate: %.1f /100k"),
    dataset$County, dataset[,ldi_feature[1]], (dataset[,ldi_feature[2]])*100000
  ) %>% lapply(htmltools::HTML)
  

  return (leaflet(shapes) %>%
  addPolygons(
    fillColor = ~pal2(dataset[,ldi_feature[1]]),
    weight = 1,
    opacity = 1,
    color = "#330000",
    dashArray = "1",
    fillOpacity = 0.7,
    highlight = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>% 
  addLegend(pal = pal2, 
            values = ~dataset[ldi_feature[1]], 
            opacity = 0.7, 
            title = paste0("Disparity Index<br/>",state.choice," COVID-19 ",feature," Rates"),
            position = "bottomright",
            labFormat = function(type, cuts, p) { n = length(cuts) 
            cuts[n] = paste0(cuts[n]," lower") 
            # for (i in c(1,seq(3,(n-1)))){cuts[i] = paste0(cuts[i],"—")} 
            for (i in c(1,seq(2,(n-1)))){cuts[i] = paste0(cuts[i]," — ")} 
            cuts[2] = paste0(cuts[2]," higher") 
            paste0(str_remove(cuts[-n],"higher"), str_remove(cuts[-1],"—"))
            }
  ) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light"))
    )
}

```

covid_confirmed_usafacts metadata:
+ countyFIPS: 0 for statewide unallocated, 1 for NYC unallocated, otherwise>1000
+ County Name: Format is "x County", i.e. "Albany County"
+ State: 2 letter initial
+ stateFIPS: 1-56
+ 1/22/20..MostRecentDate: Time series case data (aggregate)

NY.data (to replicate):
+ 10 NY.shape columns (inner join on county fips?)
+ County: County Name without 'County' in strings
+ Population
+ Date
+ Cases/Case Rate/LDI (Basically any data I want to show)

Population
+ countyFIPS: 0 for statewide unallocated, 1 for NYC unallocated, otherwise>1000
+ County Name: Format is "x County", i.e. "Albany County"
+ State: 2 letter initial
+ Population: of county

covid_deaths_usafacts metadata:
+ countyFIPS: 0 for statewide unallocated, 1 for NYC unallocated, otherwise>1000
+ County Name: Format is "x County", i.e. "Albany County"
+ State: 2 letter initial
+ stateFIPS: 1-56
+ 1/22/20..MostRecentDate: Time series case data (aggregate)


```{r data processing}
# Read in data
todays.raw.case.data <- read_csv("data/csv/time_series/covid_confirmed_usafacts.csv") 
todays.raw.death.data <- read_csv("data/csv/time_series/covid_deaths_usafacts.csv")
population <- read_csv("data/csv/time_series/covid_county_population_usafacts.csv") %>%
  filter(countyFIPS > 1000)
diabetes_data_counties <- read_csv("Summer_2020_notebooks/DiabetesAtlasCountyData.csv")
US.diabetes.pt <- 8.5

obesity_data_counties <- read_csv("Summer_2020_notebooks/adultobesity_uscounties.csv")
US.obesity.pt <- pUS.8.obesity #31.3


todays.case.data <- todays.raw.case.data %>%
  filter(countyFIPS > 1000) %>%
  select(countyFIPS, `County Name`, State, ncol(todays.raw.case.data)) %>%
  rename(c("County" = "County Name"))
colnames(todays.case.data)[4] <- "Cases"


todays.case.data <- inner_join(todays.case.data, population[c(1,4)], by=c("countyFIPS" = "countyFIPS")) %>%
  filter(population > 0)

todays.death.data <- todays.raw.death.data %>%
  filter(countyFIPS > 1000) %>%
  select(1, ncol(todays.raw.death.data))
colnames(todays.death.data)[2] <- "Mortality"

todays.case.data <- inner_join(todays.case.data, todays.death.data, by=c("countyFIPS" = "countyFIPS"))

todays.case.data$Case_rate <- todays.case.data$Cases/todays.case.data$population
todays.case.data$Case_rate_ldi <- unlist(lapply(todays.case.data$Case_rate, FUN=function(x){-log(pUS.6.cases/x)})) 
todays.case.data <- todays.case.data %>%
 mutate(Case_rate_ldi = replace(Case_rate_ldi, Case_rate_ldi < -5, -5))

todays.case.data$Mortality_rate <- todays.case.data$Mortality/todays.case.data$population
todays.case.data$Mortality_rate_ldi <- unlist(lapply(todays.case.data$Mortality_rate, FUN=function(x){-log(pUS.6.deaths/x)}))
todays.case.data <- todays.case.data %>%
 mutate(Mortality_rate_ldi = replace(Mortality_rate_ldi, Mortality_rate_ldi < -5, -5))

todays.case.data <- inner_join(todays.case.data, 
                              diabetes_data_counties[c("CountyFIPS","Percentage")], 
                              by=c("countyFIPS" = "CountyFIPS")) %>%
  rename(Diabetes_rate = Percentage)
todays.case.data$Diabetes_rate <- as.numeric(todays.case.data$Diabetes_rate)/100 # Percentage
todays.case.data$Diabetes_rate_ldi <- unlist(lapply(todays.case.data$Diabetes_rate, FUN=function(x){-log(US.diabetes.pt/(x*100))}))
todays.case.data <- todays.case.data %>%
 mutate(Diabetes_rate_ldi = replace(Diabetes_rate_ldi, Diabetes_rate_ldi < -5, -5))

obesity_data_counties$FIPS <- as.numeric(obesity_data_counties$FIPS)
todays.case.data <- inner_join(todays.case.data,
                               obesity_data_counties[c("FIPS", "% Adults with Obesity")] ,
                               by=c("countyFIPS" = "FIPS")) %>%
  rename(Obesity_rate = `% Adults with Obesity`)
todays.case.data$Obesity_rate <- as.numeric(todays.case.data$Obesity_rate)/100 # Percentage
todays.case.data$Obesity_rate_ldi <- unlist(lapply(todays.case.data$Obesity_rate, FUN=function(x){-log(US.obesity.pt/(x*100))}))
todays.case.data <- todays.case.data %>%
 mutate(Obesity_rate_ldi = replace(Obesity_rate_ldi, Obesity_rate_ldi < -5, -5))


todays.case.data <- todays.case.data %>%
  mutate(County = str_remove_all(County, regex(" County", ignore_case = T)))

```

```{r Cases}
selected_state <- "US"
selected_feature <- "Obesity"
geo.plot(selected_state, selected_feature)
```

```{r Mortality}
#selected_state <- "AL"
selected_feature <- "Mortality"
geo.plot(selected_state, selected_feature)
```

```{r State policy processing}
state_policy <- read_csv("Summer_2020_notebooks/state_policy.csv")
state_policy.df <- state_policy[5:55,]
state_policy.md <- state_policy[1:4,]
state_policy.df
#as.Date(state_policy.df$STAYHOME, format = "%m/%d/%y")
state_policy.df[as.vector(state_policy.md[4,] == "date")] <- lapply(state_policy.df[as.vector(state_policy.md[4,] == "date")], function(x){parse_date_time(x, c("%m/%d/%y"))})
```

```{r plot processing}
### County Level ###
covid_TS_counties_long.cases <- todays.raw.case.data %>%
  tidyr::gather(date,cases,5:ncol(todays.raw.case.data))
covid_TS_counties_long.death <- todays.raw.death.data %>%
  tidyr::gather(date,deaths,5:ncol(todays.raw.death.data))

# Make date column an actual R date_time
covid_TS_counties_long.cases$date <- parse_date_time(covid_TS_counties_long.cases$date, c("%m/%d/%y"))
covid_TS_counties_long.death$date <- parse_date_time(covid_TS_counties_long.death$date, c("%m/%d/%y"))

# Factor does not make sense here as there may be distint, same named counties
covid_TS_counties_long.cases <- covid_TS_counties_long.cases %>%
  rename(County = `County Name`) %>%
  mutate(County = str_remove_all(County, regex(" County", ignore_case = T)))

# Join death data
covid_TS_counties_long.cases <- inner_join(covid_TS_counties_long.cases, covid_TS_counties_long.death[c(1,5,6)], by=c("countyFIPS" = "countyFIPS", "date" = "date"))

# Filter small data
covid_TS_counties_long.cases <- covid_TS_counties_long.cases %>%
  filter(countyFIPS > 1000) %>%
  filter(cases > 5)

# Join population data
covid_TS_counties_long.cases <- left_join(covid_TS_counties_long.cases, population[c(1,4)], by=c("countyFIPS" = "countyFIPS"))

# Per100k columns
covid_TS_counties_long.cases$p_cases <- covid_TS_counties_long.cases$cases/covid_TS_counties_long.cases$population*100000

covid_TS_counties_long.cases$p_deaths <- covid_TS_counties_long.cases$deaths/covid_TS_counties_long.cases$population*100000

# Changes in cases
covid_TS_counties_long.cases %>% group_by(County, State) %>% 
  mutate(diff = ifelse(as.Date(date - 1) == lag(date), cases - lag(cases), cases)) -> 
  covid_TS_counties_long.cases

covid_TS_counties_long.cases %>% 
  mutate(p_diff = ifelse(as.Date(date - 1) == lag(date), p_cases - lag(p_cases), p_cases)) %>%
  ungroup() -> covid_TS_counties_long.cases

covid_TS_counties_long.cases$diff <- ifelse(is.na(covid_TS_counties_long.cases$diff), covid_TS_counties_long.cases$cases, covid_TS_counties_long.cases$diff)
covid_TS_counties_long.cases$p_diff <- ifelse(is.na(covid_TS_counties_long.cases$p_diff), covid_TS_counties_long.cases$p_cases, covid_TS_counties_long.cases$p_diff)

# Changes in deaths
covid_TS_counties_long.cases %>% group_by(County, State) %>% 
  mutate(d_diff = ifelse(as.Date(date - 1) == lag(date), deaths - lag(deaths), deaths)) -> 
  covid_TS_counties_long.cases

covid_TS_counties_long.cases %>% 
  mutate(p.d_diff = ifelse(as.Date(date - 1) == lag(date), p_deaths - lag(p_deaths), p_deaths)) %>%
  ungroup() -> covid_TS_counties_long.cases

covid_TS_counties_long.cases$d_diff <- ifelse(is.na(covid_TS_counties_long.cases$d_diff), covid_TS_counties_long.cases$deaths, covid_TS_counties_long.cases$d_diff)
covid_TS_counties_long.cases$p.d_diff <- ifelse(is.na(covid_TS_counties_long.cases$p.d_diff), covid_TS_counties_long.cases$p_deaths, covid_TS_counties_long.cases$p.d_diff)

### State Level ###

covid_TS_state_long.cases <- todays.raw.case.data %>%
  tidyr::gather(date,cases,5:ncol(todays.raw.case.data))
covid_TS_state_long.death <- todays.raw.death.data %>%
  tidyr::gather(date,deaths,5:ncol(todays.raw.death.data))

# Make date column an actual R date_time
covid_TS_state_long.cases$date <- parse_date_time(covid_TS_state_long.cases$date, c("%m/%d/%y"))
covid_TS_state_long.death$date <- parse_date_time(covid_TS_state_long.death$date, c("%m/%d/%y"))

# Factor does not make sense here as there may be distint, same named counties
covid_TS_state_long.cases <- covid_TS_state_long.cases %>%
  group_by(State, date) %>%
  summarise(cases = sum(cases))

covid_TS_state_long.death <- covid_TS_state_long.death %>%
  group_by(State, date) %>%
  summarise(deaths = sum(deaths))

# Join death data
covid_TS_state_long.cases <- inner_join(
  covid_TS_state_long.cases, covid_TS_state_long.death, by=c("State" = "State", "date" = "date"))

# Filter small data
covid_TS_state_long.cases <- covid_TS_state_long.cases %>%
  filter(cases > 5)

# Join population data
covid_TS_state_long.cases <- left_join(
  covid_TS_state_long.cases, 
  population %>% group_by(State) %>% summarise(population = sum(population)), 
  by=c("State" = "State"))

# Per100k columns
covid_TS_state_long.cases$p_cases <- covid_TS_state_long.cases$cases/covid_TS_state_long.cases$population*100000

covid_TS_state_long.cases$p_deaths <- covid_TS_state_long.cases$deaths/covid_TS_state_long.cases$population*100000

# Changes in cases
covid_TS_state_long.cases %>% group_by(State) %>% 
  mutate(diff = ifelse(as.Date(date - 1) == lag(date), cases - lag(cases), cases)) -> 
  covid_TS_state_long.cases

covid_TS_state_long.cases %>% 
  mutate(p_diff = ifelse(as.Date(date - 1) == lag(date), p_cases - lag(p_cases), p_cases)) %>%
  ungroup() -> covid_TS_state_long.cases

covid_TS_state_long.cases$diff <- ifelse(is.na(covid_TS_state_long.cases$diff), covid_TS_state_long.cases$cases, covid_TS_state_long.cases$diff)
covid_TS_state_long.cases$p_diff <- ifelse(is.na(covid_TS_state_long.cases$p_diff), covid_TS_state_long.cases$p_cases, covid_TS_state_long.cases$p_diff)

# Changes in deaths
covid_TS_state_long.cases %>% group_by(State) %>% 
  mutate(d_diff = ifelse(as.Date(date - 1) == lag(date), deaths - lag(deaths), deaths)) -> 
  covid_TS_state_long.cases

covid_TS_state_long.cases %>% 
  mutate(p.d_diff = ifelse(as.Date(date - 1) == lag(date), p_deaths - lag(p_deaths), p_deaths)) %>%
  ungroup() -> covid_TS_state_long.cases

covid_TS_state_long.cases$d_diff <- ifelse(is.na(covid_TS_state_long.cases$d_diff), covid_TS_state_long.cases$deaths, covid_TS_state_long.cases$d_diff)
covid_TS_state_long.cases$p.d_diff <- ifelse(is.na(covid_TS_state_long.cases$p.d_diff), covid_TS_state_long.cases$p_deaths, covid_TS_state_long.cases$p.d_diff)


# State ranking
time.period <- 14
ranking <- covid_TS_state_long.cases %>%
  group_by(State) %>%
  filter(State != "DC") %>%
  top_n(time.period, wt=date) %>%
  summarise(
    population = max(population),
    cases.delta = (max(p_cases) - min(p_cases))/max(p_cases),
    deaths.delta =  (max(p_deaths) - min(p_deaths))/max(p_deaths)
    ) %>%
  #mutate(points = 0.5*(cases.delta + deaths.delta)) %>%
  mutate(points = cases.delta) %>%
  arrange(points) %>%
  mutate(rank = row_number()) %>%
  left_join(state.abr[c("abr", "name")],
            by = c("State" = "abr"))

### US Level ###

covid_TS_US_long.cases <- todays.raw.case.data %>%
  tidyr::gather(date,cases,5:ncol(todays.raw.case.data))
covid_TS_US_long.death <- todays.raw.death.data %>%
  tidyr::gather(date,deaths,5:ncol(todays.raw.death.data))

# Make date column an actual R date_time
covid_TS_US_long.cases$date <- parse_date_time(covid_TS_US_long.cases$date, c("%m/%d/%y"))
covid_TS_US_long.death$date <- parse_date_time(covid_TS_US_long.death$date, c("%m/%d/%y"))

# Factor does not make sense here as there may be distint, same named counties
covid_TS_US_long.cases <- covid_TS_US_long.cases %>%
  group_by(date) %>%
  summarise(cases = sum(cases))

covid_TS_US_long.death <- covid_TS_US_long.death %>%
  group_by(date) %>%
  summarise(deaths = sum(deaths))

# Join death data
covid_TS_US_long.cases <- inner_join(
  covid_TS_US_long.cases, covid_TS_US_long.death, by=c("date" = "date"))

# Filter small data
covid_TS_US_long.cases <- covid_TS_US_long.cases %>%
  filter(cases > 5)

# Join population data
US.pop <- sum(population$population)

# Per100k columns
covid_TS_US_long.cases$p_cases <- covid_TS_US_long.cases$cases/US.pop*100000

covid_TS_US_long.cases$p_deaths <- covid_TS_US_long.cases$deaths/US.pop*100000

# Changes in cases
covid_TS_US_long.cases %>%
  mutate(diff = ifelse(as.Date(date - 1) == lag(date), cases - lag(cases), cases)) -> 
  covid_TS_US_long.cases

covid_TS_US_long.cases %>% 
  mutate(p_diff = ifelse(as.Date(date - 1) == lag(date), p_cases - lag(p_cases), p_cases)) -> 
  covid_TS_US_long.cases

covid_TS_US_long.cases$diff <- ifelse(is.na(covid_TS_US_long.cases$diff), covid_TS_US_long.cases$cases, covid_TS_US_long.cases$diff)
covid_TS_US_long.cases$p_diff <- ifelse(is.na(covid_TS_US_long.cases$p_diff), covid_TS_US_long.cases$p_cases, covid_TS_US_long.cases$p_diff)

# Changes in deaths
covid_TS_US_long.cases %>% 
  mutate(d_diff = ifelse(as.Date(date - 1) == lag(date), deaths - lag(deaths), deaths)) -> 
  covid_TS_US_long.cases

covid_TS_US_long.cases %>% 
  mutate(p.d_diff = ifelse(as.Date(date - 1) == lag(date), p_deaths - lag(p_deaths), p_deaths)) -> 
  covid_TS_US_long.cases

covid_TS_US_long.cases$d_diff <- ifelse(is.na(covid_TS_US_long.cases$d_diff), covid_TS_US_long.cases$deaths, covid_TS_US_long.cases$d_diff)
covid_TS_US_long.cases$p.d_diff <- ifelse(is.na(covid_TS_US_long.cases$p.d_diff), covid_TS_US_long.cases$p_deaths, covid_TS_US_long.cases$p.d_diff)
```


```{r get_y_label}
get_y_label <- function(y.value) {
  if (y.value == "cases") {
    return("COVID-19 Cases")
  }
  if (y.value == "p_cases") {
    return("COVID-19 Cases per 100k")
  }
  if (y.value == "diff") {
    return("New COVID-19 Cases")
  }
  if (y.value == "p_diff") {
    return("New COVID-19 Cases per 100k")
  }
  if (y.value == "deaths") {
    return("COVID-19 Deaths")
  }
  if (y.value == "p_deaths") {
    return("COVID-19 Deaths per 100k")
  }
  else {
    return("")
  }
}
```

```{r ggplot.state}
ggplot.state <- function(selected.state="NY", y.value="cases", moving.avg.window=7, case.cut=200, max.labels=10, remove.title = F) {
  
  y_label <- get_y_label(y.value)
  state.name <- state.abr[state.abr$abr==selected.state,"name"]
  m.a.w <- ""
  if (y.value == "diff" | y.value == "p_diff") {
    m.a.w <- paste0(" (",moving.avg.window," day Average)")
  }
  
  if (remove.title) {
    gg_title <- NULL
  }
  else {
    gg_title <- ggtitle(paste0(state.name, " ", y_label, " over time",m.a.w))
  }
  
  
  covid_TS_counties.cases.plot <- covid_TS_counties_long.cases %>%
    filter(State == selected.state) %>%
    group_by(County) %>% 
    filter(n() >= moving.avg.window) %>%
    mutate(diff = c(numeric(moving.avg.window-1), zoo::rollmean(diff, moving.avg.window, align = "right"))) %>%
    mutate(p_diff = c(numeric(moving.avg.window-1), zoo::rollmean(p_diff, moving.avg.window, align = "right"))) %>%
    ungroup()
  
  if (selected.state == "NY") {
    nyc.population <- covid_TS_counties.cases.plot %>%
      filter(County %in% c("New York", "Kings", "Queens", "Bronx", "Richmond")) %>%
      group_by(County) %>%
      top_n(n=1, wt=date) %>%
      select(population)
    nyc.population <- sum(nyc.population$population)
      
    NYC <- covid_TS_counties.cases.plot %>%
      filter(County %in% c("New York", "Kings", "Queens", "Bronx", "Richmond")) %>%
      group_by(State, date, stateFIPS) %>%
      summarise(
        countyFIPS = 1,
        County = "New York City",
        cases = sum(cases),
        deaths = sum(deaths),
        population = nyc.population,
        p_cases = sum(cases)*100000/nyc.population,
        p_deaths = sum(deaths)*100000/nyc.population,
        diff = sum(diff),
        p_diff = sum(diff)*100000/nyc.population,
        d_diff = sum(d_diff),
        p.d_diff = sum(p.d_diff)*100000/nyc.population
      ) 
    covid_TS_counties.cases.plot <- covid_TS_counties.cases.plot %>%
      filter(!County %in% c("New York", "Kings", "Queens", "Bronx", "Richmond")) %>%
      rbind.data.frame(NYC)
      
  }
  
  state_pop <- ranking[ranking$State == selected.state,]$population

  state <- covid_TS_counties.cases.plot %>%
    group_by(State, date, stateFIPS) %>%
    summarise(
      countyFIPS = 0,
      County = selected.state,
      cases = sum(cases),
      deaths = sum(deaths),
      population = state_pop,
      p_cases = sum(cases)*100000/state_pop,
      p_deaths = sum(deaths)*100000/state_pop,
      diff = sum(diff),
      p_diff = sum(diff)*100000/state_pop,
      d_diff = sum(d_diff),
      p.d_diff = sum(p.d_diff)*100000/state_pop
    )
  
  n_county <- covid_TS_counties.cases.plot$County %>% unique() %>% length()
  covid_TS_counties.cases.plot <-  covid_TS_counties.cases.plot %>%
    group_by(County) %>%
    filter(max(cases) > case.cut) %>%
    ungroup() %>%
    rbind.data.frame(state) %>%
    filter(get(y.value) > 0)
    
  
  county.num <- covid_TS_counties.cases.plot %>% 
        select(County) %>%
        unique()
  
  colnames(county.num) <- c("County")
  county.num <- county.num %>%
    mutate(num = row_number()) %>%
    mutate(County = as.character(County))
  
  n_county <- nrow(county.num)
  highlight_points <- covid_TS_counties.cases.plot  %>%
    left_join(county.num, by=c("County" = "County")) %>%
    group_by(County) %>%
    mutate(range = as.numeric(as.Date(max(date))) - as.numeric(as.Date(min(date)))) %>%
    mutate(max_date = as.Date(max(date))) %>%
    mutate(max_date = max_date - ((num*((range%/%n_county) + 1))%%range))  %>%
    filter(date == max_date) %>%
    top_n(1, wt=max_date) 
  
  library(RColorBrewer)
  qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
  col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
  county.num$Color <- col_vector[1:n_county]
  
  region_palette <- setNames(as.character(county.num$Color), as.character(county.num$County))
  region_palette[selected.state] <- "#A8A8A8"

  g <- covid_TS_counties.cases.plot %>%
          ggplot(aes_string(
                   x="date",
                   y=y.value,
                   color="County",
                   group="County")) +
                scale_color_manual(values=region_palette, guide=guide_legend(title.position = "top",title.hjust = 0.5)) +
                geom_line(size=1) +
                scale_y_continuous(
                  trans = "log10",
                  breaks = c(10,100,500,1000,5000,10000, 50000)
                ) +
                scale_x_datetime(date_breaks = "1 week", date_minor_breaks = "1 day", date_labels = "%b %d") +
                ylab(y_label) + 
                xlab("Date") +
                #theme(legend.position = "none") +
                geom_label_repel(
                  data=highlight_points,  
                  aes(label=County, fill=County), 
                  box.padding = unit(1.75, 'lines'),
                  color = "black",
                  size = 5,
                  show.legend = FALSE
                ) +
                scale_color_manual(values=region_palette, aesthetics = c("fill")) +
                geom_vline(aes(xintercept=state_policy.df[state_policy.df$POSTCODE == selected.state,]$END_STHM, linetype=paste0(selected.state," ends stay at home order")), color = "red") + 
                geom_vline(aes(xintercept=state_policy.df[state_policy.df$POSTCODE == selected.state,]$STAYHOME, linetype=paste0(selected.state," begins stay at home order")), color = "blue") + 
                scale_linetype_manual(name = "Events", 
                                      values = c(2,2), 
                                      guide = guide_legend(title.position = "top",title.hjust = 0.5,override.aes = list(color = c("blue", "red")), direction = "vertical")) +
                theme(legend.position = "bottom") +
                gg_title +
                NULL
    return(g)
}
```

```{r}
selected_state <- "WV"
moving.avg.window <- 7
n_case_cut <- 100
y.value <- "diff"
ggplot.state(selected_state, y.value=y.value, case.cut=n_case_cut, moving.avg.window=moving.avg.window)
```

```{r get_dif (label)}
get_dif <- function(y.value) {
  if (y.value == "cases") {
    return("diff")
  }
  if (y.value == "deaths") {
    return("d_diff")
  }
  if (y.value == "p_cases") {
    return("p_diff")
  }
  if (y.value == "p_deaths") {
    return("p.d_diff")
  }
}
```

```{r Case bar chart}
ggbar.overall <- function(selected.state = "VT", y.value="p_deaths", moving.avg.window=14, remove.title = F) {
  state <- covid_TS_state_long.cases %>%
    filter(State == selected.state)
  
  my_diff <- get_dif(y.value)
  category <- get_y_label(y.value)
  
  #y.max <- max(covid_TS_state_long.cases[y.value])
  
  if (remove.title) {
    gg_title <- NULL
  }
  else {
    gg_title <- ggtitle(paste0(selected.state, " ", category, " over time"))
  }
  
  
  state_cases <- state[c("date", y.value, my_diff)]
  state_cases <- state_cases %>%
    rename(Values = all_of(y.value)) %>%
    rename(diff = all_of(my_diff)) %>%
    mutate(diff.ma =  c(diff[1:7-1], zoo::rollmean(diff, 7, align="right"))) %>%
    mutate(pct_increase =diff.ma/Values*100) %>%
    mutate(ma = c(numeric(moving.avg.window-1), zoo::rollmean(Values, moving.avg.window, align = "right")))
  state_cases[state_cases$diff.ma > 0 & state_cases$pct_increase > 10, "pct_increase"] <- 10
  state_cases[is.na(state_cases$pct_increase) | state_cases$pct_increase < 0.01, "pct_increase"] <- NA
  state_cases$Type <- "State MA"
  
  US.ma <- covid_TS_US_long.cases[c("date", y.value, my_diff)] %>%
    rename(Values = all_of(y.value)) %>%
    rename(diff = all_of(my_diff)) %>%
    mutate(diff.ma =  c(diff[1:7-1], zoo::rollmean(diff, 7, align="right"))) %>%
    mutate(pct_increase =diff.ma/Values*100) %>%
    mutate(ma = c(numeric(moving.avg.window-1), zoo::rollmean(Values, moving.avg.window, align = "right"))) %>%
    filter(ma > 0) %>%
    filter(date > min(state_cases$date))
  US.ma$Type <- "US MA"
  state.us.ma <- rbind.data.frame(state_cases, US.ma)
  
  END_STHM <- paste0(selected.state," ends stay at home order")
  STHM <- paste0(selected.state," begins stay at home order")
  
  return(state_cases %>%
    ggplot() +
    geom_col(aes(x=date, y=Values, fill = pct_increase)) +
    scale_fill_gradient(high = "#ff0000", 
                        low = "#ffffff",
                        limit = c(0,10),
                        breaks = c(5,10),
                        labels = c("5%","10%+"),
                        na.value = "skyblue",
                        guide = guide_colorbar(title = paste0("Percentage change in ", category),
                                               title.hjust = 0.5,
                                               title.position = "top", 
                                               label.hjust = 0.5, 
                                               barwidth = 15,
                                               frame.colour = "black")) +
    geom_point(data = state_cases, aes(x=date, y=Values, size = ""), shape = NA, colour = "skyblue") +
    guides(size=guide_legend(title = "No Change", 
                             override.aes=list(shape=15, size = 8), 
                             title.position = "top",
                             title.hjust = 0.5)) +
    geom_line(data = state.us.ma,aes( x=date, y=ma, color=Type, group=Type), arrow=arrow(ends="last"), show.legend = F)  + 
    geom_line(data = state.us.ma,aes( x=date, y=ma, color=Type, group=Type)) +
    geom_vline(aes(xintercept=state_policy.df[state_policy.df$POSTCODE == selected.state,]$END_STHM, color = END_STHM), linetype="longdash", size = 1, show.legend = F) +
    geom_vline(aes(xintercept=state_policy.df[state_policy.df$POSTCODE == selected.state,]$STAYHOME, color= STHM), linetype="longdash", size = 1, show.legend = F) +
    scale_color_manual(name = "Line Types", 
                        values = c("blue", "red", "black", "steelblue"),
                        guide = guide_legend(title.position = "top",
                                             title.hjust = 0.5,
                                             direction = "vertical")) +
    scale_x_datetime(date_breaks = "1 week", date_minor_breaks = "1 day", date_labels = "%b %d") +
    ylab(get_y_label(y.value)) + 
    xlab("Date") +
    theme(legend.position = "bottom",legend.direction = "horizontal") +
    gg_title + 
    NULL
  )
}
                
```


```{r}
ggbar.overall("HI", "p_deaths")
```

```{r}
ma.plot <- function(selected.state = "AZ", y.value="p_deaths", moving.avg.window=14, remove.title = F) {
  state <- covid_TS_state_long.cases %>%
    filter(State == selected.state)
  
  category <- get_y_label(y.value)
  
  #y.max <- max(covid_TS_state_long.cases[y.value])
  
  if (remove.title) {
    gg_title <- NULL
  }
  else {
    gg_title <- ggtitle(paste0(selected.state, 
                               " ", 
                               category, 
                               " compared to US over time (",
                               moving.avg.window,
                               " day average)"))
  }
  
  state_cases <- state[c("date", y.value)]
  state_cases <- state_cases %>%
    mutate(ma = c(numeric(moving.avg.window-1), zoo::rollmean(eval(parse(text=y.value)), moving.avg.window, align = "right"))) %>%
    filter(ma > 0) %>%
    select(date, ma)
  state_cases$Type = "State MA"
  
  US.ma <- covid_TS_US_long.cases[c("date", y.value)] %>%
    mutate(ma = c(numeric(moving.avg.window-1), zoo::rollmean(eval(parse(text=y.value)), moving.avg.window, align = "right"))) %>%
    filter(ma > 0) %>%
    select(date, ma)
  US.ma$Type <- "US MA"
  state_cases <- rbind.data.frame(state_cases, US.ma)
  END_STHM <- paste0(selected.state," ends stay at home order")
  STHM <- paste0(selected.state," begins stay at home order")
  return(
    state_cases %>%
           ggplot(aes(x = date, 
                      y = ma, 
                      color = Type,
                      group = Type)) +
    geom_line(arrow = arrow(ends = "last"), show.legend = F) +
    geom_line() +
    geom_vline(aes(xintercept=state_policy.df[state_policy.df$POSTCODE == selected.state,]$END_STHM, color = END_STHM), linetype="dashed", size=1, show.legend = F) +
    geom_vline(aes(xintercept=state_policy.df[state_policy.df$POSTCODE == selected.state,]$STAYHOME, color= STHM), linetype="dashed", size = 1, show.legend = F) +
    scale_color_manual(name = "Events", 
                        values = c("black", "red", "green", "steelblue"),
                        guide = guide_legend(title.position = "top",
                                             title.hjust = 0.5,
                                             direction = "vertical")) +
      NULL
  )
}
```


```{r gt.ranking}
gt.ranking <- function(entries = 50, order=function(x){x}) {
  ranking %>%
    arrange(order(rank)) %>%
    filter(row_number() <= entries) %>%
    select(name, cases.delta, deaths.delta, rank) %>%
    gt() %>%
    tab_header(
      title = paste0("State Rankings"),
      subtitle = "Based on 14 day change in COVID-19 Cases and Deaths"
    ) %>%
    tab_source_note(
      source_note = md("Data Source: [USA Facts](https://usafacts.org/visualizations/coronavirus-covid-19-spread-map/)")
    ) %>%
    fmt_percent(c("cases.delta", "deaths.delta"), decimals = 1) %>%
    cols_label(
      name = "State",
      cases.delta = "Change in Cases",
      deaths.delta = "Change in Deaths",
      rank = "Rank"
    ) %>%
    cols_move_to_start(columns = vars(rank)) %>%
    tab_options(container.width = "100%",
                table.width = "100%")
}
gt.ranking(10,dplyr::desc)
```

```{r State stats chart}
#New York Stats
# tags$b("Overall Cases: "),format(stats$cases, big.mark = ","), tags$br(),
#        tags$b("Cases per 100k: "),format(round(stats$p_cases), big.mark = ","), tags$br(),
#        tags$b("Overall Deaths: "),format(stats$deaths, big.mark = ","), tags$br(),
#        tags$b("Deaths per 100k: "),format(round(stats$p_deaths), big.mark = ","), tags$br(),
#        tags$b("Stay At Home Order Start Date: "), format(policy$STAYHOME, "%m-%d-%Y"), tags$br(),
#        tags$b("Stay At Home Order End Date: "), format(policy$END_STHM, "%m-%d-%Y")

selected_state <- "NY"
numeric_fmt <- function(.x) {
  num.x <- as.numeric(.x)
  str.x <- format(round(num.x), big.mark = ",")
  str.x
}

date_fmt <- function(.x) {
  date.x <- as.Date(.x)
  str.x <- format(date.x, "%m-%d-%Y")
  str.x
}

sup_fmt <- function(.x, sup=1) {
  md(glue::glue("{.x}<sup>{sup}</sup>"))
}

sup2_fmt <- function(.x) {
  sup_fmt(.x, sup=2)
}

sup3_fmt <- function(.x) {
  sup_fmt(.x, sup=3)
}

stats.table <- function(selected_state) {
  # TODO: Instead of row numbers, add aditional column specifying data type
  state_name <- paste0(state.abr[state.abr$abr == selected_state, "name"])
  
  stats <- covid_TS_state_long.cases %>%
    filter(State == selected_state) %>%
    top_n(n=1, wt=date) %>%
    select(cases, p_cases, deaths, p_deaths) %>%
    t()
  
  tests <- state_covid_testing %>%
    filter(NAME == state_name) %>%
    select(total_num_tests, tests_per_1000) %>%
    mutate(tests_per_1000 = tests_per_1000 * 100) %>%
    t()
  
  policy <- state_policy.df %>%
    filter(POSTCODE == selected_state) %>% 
    select(STAYHOME, END_STHM) %>%
    t()
  
  stats <- stats %>%
    rbind.data.frame(tests, policy)

  state.title <- paste0(state_name, " Stats")
  stats$features <- c("Overall Cases",  "Overall Cases per 100k", "Overall Deaths", "Overall Deaths per 100k", "Overall Tests", "Overall Tests per 100k", "Stay At Home Order Start Date", "Stay At Home Order End Date")
  
  stats %>%
    mutate(row_num = row_number()) %>%
    gt() %>%
    fmt(c("features"), row = row_num < 5, fns = sup_fmt) %>%
    fmt(c("features"), row = row_num >= 5 & row_num < 7, fns = sup2_fmt) %>%
    fmt(c("features"), row = row_num >= 7, fns = sup3_fmt) %>%
    cols_move_to_start(c("features")) %>%
    tab_header(
      title = md(paste0("**", state.title, "**"))
    ) %>%
    tab_source_note(
        source_note = md("Data Source: [USA Facts<sup>1</sup>](https://usafacts.org/visualizations/coronavirus-covid-19-spread-map/), [OWID<sup>2</sup>](https://covid.ourworldindata.org/data/owid-covid-data.csv)")
      ) %>%
    tab_source_note(
        source_note = md("Data Source: Raifman J, Nocka K, Jones D, Bor J, Lipson S, Jay J, and Chan P. (2020). 'COVID-19 US state policy database.' Available at: [www.tinyurl.com/statepolicies<sup>3</sup>](www.tinyurl.com/statepolicies)")
      ) %>%
    fmt(c("V1"), row = row_num < 7, fns = numeric_fmt) %>%
    fmt(c("V1"), row = row_num >= 7, fns = date_fmt) %>%
    cols_hide(c("row_num")) %>%
    tab_options(column_labels.hidden = T,
                container.width = "100%",
                table.width = "100%")
}
stats.table("HI")
```


```{r shinyApp}
shinyApp(
  ui = fluidPage(
    includeCSS("www/style.css"),
    tags$style(HTML(".selectize-control.single .selectize-input {
    width: 100%;
    max-height: 50px;
}")),
    fluidPage(
      fluidRow(column(12, style="text-align:center;",
                      selectInput(inputId = "state_name",
                                  label = "State Selector",
                                  choices = state.abr$name,
                                  selected = as.character(unlist(ranking[ranking$rank==50, "name"]))))),
      fluidRow(column(12, style="text-align:center;",uiOutput("main_title"))),
      tags$br(),
      fluidRow(column(8, style="text-align:center;",
                      tags$b(tags$sup("*"),"States are ranked best to worst by their percentage change in COVID-19 cases over the past 14 days."),
                      offset=2)),
      fluidRow(column(6, style="text-align:center;",uiOutput("state.CoT.title"),
                      plotOutput(outputId = "state.CoT", height = "600px")),
               column(6, style="text-align:center;",uiOutput("state.DoT.title"),
                      plotOutput(outputId = "state.DoT", height = "600px"))),
      fluidRow(column(8, style="text-align:center;",
                      tags$h2("Flattening the Curve"),
                      tags$p("Nationwide, states have taken various approaches to mitigate the spread of coronavirus, such as social distancing interventions and encouraging mask use where social distancing is not possible. Studies by the CDC have shown these methods reduce new COVID-19 cases, hospitalizations, and deaths."),
                      tags$b("Data Source: "), tags$a("CDC", href="https://wwwnc.cdc.gov/eid/article/26/8/20-1093_article"), offset=2)),
                      tags$br(),
                      tags$br(),
      fluidRow(column(8,gt_output("state.report"),
                      offset = 2)),
      fluidRow(column(12, style="text-align:center;",
                      tags$h1("County Level Breakdown"))),
      fluidRow(column(10, style="text-align:center;",uiOutput("state.trends.title"),
                      plotOutput(outputId = "state.trends", height="600px"), offset = 1)),
      fluidRow(column(6,
                      uiOutput("state.county.cases"),
                      leafletOutput("map.cases")),
               column(6,
                      uiOutput("state.county.deaths"),
                      leafletOutput("map.deaths"))),
      tags$br(),
      fluidRow(column(12, style="text-align:center;",
                      tags$h1("Comorbidities"))),
      fluidRow(column(12, style="text-align:center;",
                      uiOutput("determinant.title")),
               column(6,
                      selectInput(inputId = "determinant",
                                  label = NULL,
                                  choices = c("Diabetes", "Obesity"),
                                  selected = "Obesity"),
                      leafletOutput("maps.determinant"), offset = 3),
               column(8, style="text-align:center;",
                      tags$p(textOutput("determinant.text"),
                             tags$br(),
                             tags$b("Data Source: "), tags$a("CDC", href="www.cdc.gov/diabetes/data"), ", ",
                             tags$a("CHR", href = 
                                      "https://www.countyhealthrankings.org/explore-health-rankings/measures-data-sources/county-health-rankings-model/health-factors/health-behaviors/diet-exercise/adult-obesity")), 
                      offset = 2
                      )),
      fluidRow(column(8, style="text-align:center;",
                       HTML("Respective rates per 100k people on maps are: 
                        <div>
                       <div>&nbsp;&nbsp;&nbsp;<span style='background: #BD0026; border-radius: 50%; font-size: 11px; opacity: 0.7;'>&nbsp&nbsp&nbsp&nbsp</span><strong> Higher</strong> than US avg. rate for disparity index &gt; 0.2</div>
                       <div>&nbsp;&nbsp;&nbsp;<span style='background: #f7f7f7; border-radius: 50%; font-size: 11px; opacity: 0.7;'>&nbsp&nbsp&nbsp&nbsp</span><strong> About equal</strong> to US avg. rate for -0.2 &lt; disparity index &lt; 0.2</div>
                       <div>&nbsp;&nbsp;&nbsp;<span style='background: #253494; border-radius: 50%; font-size: 11px; opacity: 0.7;'>&nbsp&nbsp&nbsp&nbsp</span><strong> Lower</strong> than US avg. rate for disparity index &lt; -0.2</div>
                       <i>Darker shades indicate greater disparity.</i><br><br>
                       </div>"), offset=2)),
      tags$br(),
      tags$br()
      ,fluidRow(column(12, style="text-align:center;",
                       tags$h1("Rankings")),
                tags$a(name = "ranking"),
                column(8,
                       fluidRow(style="float:right;width:250px;",
                                selectInput(inputId = "entries",
                                  label = "Entries",
                                  choices = c(`Show 10` = 10, 
                                              `Show 25` = 25,
                                              `Show 50` = 50),
                                  selected = 10,
                                  width = "50%"),
                                radioButtons(inputId = "rank.order",
                                             label = "Order",
                                             choices = c("Ascending", "Descending"),
                                             selected = "Ascending",
                                             width="50%")),
                       gt_output("ranking.table"), offset = 2))
    )
  ),
  
  
  server = function(input, output) {
    output$main_title <- renderUI({
      state_name <- input$state_name
      state_initial <- state.abr[state.abr$name == state_name, "abr"]
      tagList(
        tags$h1(paste0(state_name, " Overview")),
        tags$h2(tags$a(paste0("State rank: ", ranking[ranking$State == state_initial, "rank"]), tags$sup("*"), href = "#ranking", style="color:black;"))
      )
    })
    
    output$state.CoT.title <- renderUI({
      state_name <- input$state_name
      tags$h3(paste0(state_name, " COVID-19 Cases Over Time per 100k"))
    })
    
    output$state.DoT.title <- renderUI({
      state_name <- input$state_name
      tags$h3(paste0(state_name, " COVID-19 Deaths Over Time per 100k"))
    })
    
    output$state.trends.title <- renderUI({
      state_name <- input$state_name
      tags$h3(paste0(state_name, " ", get_y_label("diff"), " over time (7 day average)"))
    })
    
    output$state.county.cases <- renderUI({
      state_name <- input$state_name
      tags$h3(paste0(state_name, " COVID-19 Case disparities compared to US Average"))
    })
    
    output$state.county.deaths <- renderUI({
      state_name <- input$state_name
      tags$h3(paste0(state_name, " COVID-19 Mortality disparities compared to US Average"))
    })
    
    output$determinant.title <- renderUI({
      state_name <- input$state_name
      det <- input$determinant
      tags$h3(paste0(state_name, " ", det, " Rate Disparities Compared to the national average"))
    })
    
    output$determinant.text <- renderText({
      det <- input$determinant
      paste0("Nationwide, ",det," has been observed as a leading comorbidity of COVID-19.")
    })
    
    output$state.report <- render_gt({
      state_name <- input$state_name
      state_initial <- state.abr[state.abr$name == state_name, "abr"]
      stats.table(state_initial)
      })
    
    output$map.cases <- renderLeaflet({
      state_name <- input$state_name
      state_initial <- state.abr[state.abr$name == state_name, "abr"]
      geo.plot(state_initial, "Case")
      })
    
    output$map.deaths <- renderLeaflet({
      state_name <- input$state_name
      state_initial <- state.abr[state.abr$name == state_name, "abr"]
      geo.plot(state_initial, "Mortality")
      })
    
    output$maps.determinant <- renderLeaflet({
      state_name <- input$state_name
      det <- input$determinant
      state_initial <- state.abr[state.abr$name == state_name, "abr"]
      geo.plot(state_initial, det)
    })
    
    output$state.CoT <- renderPlot({
      state_name <- input$state_name
      state_initial <- state.abr[state.abr$name == state_name, "abr"]
      ggbar.overall(state_initial, y.value = "p_cases", remove.title = T)
    })
    
    output$state.DoT <- renderPlot({
      state_name <- input$state_name
      state_initial <- state.abr[state.abr$name == state_name, "abr"]
      ggbar.overall(state_initial, y.value = "p_deaths", remove.title = T)
    })
    
    output$state.trends <- renderPlot({
      state_name <- input$state_name
      state_initial <- state.abr[state.abr$name == state_name, "abr"]
      n_case_cut <- 500
      if (state_initial %in% c("MT", "WV", "AK", "HI", "SD", "WY")) {
        n_case_cut <- 100
      }
      ggplot.state(state_initial, y.value = "diff", case.cut = n_case_cut,remove.title = T)
    })
    
    output$ranking.table <- render_gt({
    entries <- input$entries
    if ("Ascending" %in% input$rank.order) {
      order <- function(x){x}
    }
    if ("Descending" %in% input$rank.order) {
      order <- dplyr::desc
    }
    gt.ranking(as.numeric(entries), order)
  })
  }
)
```




