---
title: "Using R's Epidemic Packages to Model COVID-19"
author: "Thomas Hopkins"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE}
if(!require('tidyverse')) {
  install.packages('tidyverse')
  library(tidyverse)
}
if(!require('ggplot2')) {
  install.packages('ggplot2')
  library(ggplot2)
}
if(!require('dplyr')) {
  install.packages('dplyr')
  library(dplyr)
}
if(!require('EpiEstim')) {
  install.packages('EpiEstim')
  library(EpiEstim)
}
if(!require('stringr')) {
  install.packages('stringr')
  library(stringr)
}
if(!require('matlab')) {
  install.packages('matlab')
  library(matlab)
}
knitr::opts_chunk$set(echo = TRUE)
```

Read in the time-series data for New York State. We need to calculate the incidence of COVID-19 and differentiate between imported and local cases. We assume the entries with less than 70 cases are considered imported. By imported we mean that the infected individual came from outside the region. Local cases are the entries with more than 70 cases.

```{r}
covid_TS_states <- read.csv('../data/csv/time_series/time_series_covid19_confirmed_US.csv')
covid_TS_NY <- covid_TS_states[covid_TS_states$Province_State=='New York',] %>%
  group_by(Province_State) %>%
  summarize_if(is.numeric, sum, na.rm=TRUE)
col_names_dates <- str_replace_all(str_remove(colnames(covid_TS_NY)[7:ncol(covid_TS_NY)], 'X'), coll('.'), c(pattern1 = '/'))
dates <- as.Date(col_names_dates, format = "%m/%d/%y")
covid_cases_NY <- as.numeric(covid_TS_NY[,7:ncol(covid_TS_NY)])
local_cases_NY <- covid_cases_NY[covid_cases_NY > 70]
imported_cases_NY <- covid_cases_NY[covid_cases_NY <= 70]
local_cases_NY <- as.numeric(padarray(local_cases_NY, padsize=c(0,length(dates)-length(local_cases_NY)), direction='pre'))
imported_cases_NY <- as.numeric(padarray(imported_cases_NY, padsize=c(0,length(dates)-length(imported_cases_NY)), direction='post'))
covid_cases_NY.df <- data.frame(dates=dates, local=local_cases_NY, imported=imported_cases_NY)
covid_incidence_NY.df <- covid_cases_NY.df[41:nrow(covid_cases_NY.df),]
covid_incidence_NY.df$local <- ave(covid_incidence_NY.df$local, FUN=function(x) c(0, diff(x)))
covid_incidence_NY.df$imported <- ave(covid_incidence_NY.df$imported, FUN=function(x) c(0, diff(x)))
covid_incidence_NY.df$local[6] <- covid_incidence_NY.df$local[6] - covid_cases_NY.df$imported[45]
covid_incidence_NY.df[covid_incidence_NY.df < 0] <- 0
head(covid_incidence_NY.df)
```

Now we calculate the estimated effective reproductive number ($R_e$) for COVID-19 in New York State. This is done by using the EpiEstim `estimate_R()` function. We then plot the time-series for incidence, the explored serial intervals, and $R_e$. We use a parametric serial interval curve with a mean of 3.96 and a standard deviation of 4.75. 

*Source: https://www.medrxiv.org/content/10.1101/2020.02.19.20025452v4*

```{r}
plot_Ri <- function(estimate_R_obj) {
    p_I <- plot(estimate_R_obj, "incid", add_imported_cases = TRUE)  # plots the incidence
    p_SI <- plot(estimate_R_obj, "SI")  # plots the serial interval distribution
    p_Ri <- plot(estimate_R_obj, "R")
    return(gridExtra::grid.arrange(p_I, p_SI, p_Ri, ncol = 1))
}

NY_res_parametric_si <- estimate_R(covid_incidence_NY.df, 
    method = "parametric_si", config = make_config(list(mean_si = 3.96, std_si = 4.75)))

plot_Ri(NY_res_parametric_si)
```