---
title: "MATP-4400 COVID-19 Final Notebook"
author: "Andrew Imrie"
date: "May 2020"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, results = 'hide',  warning=FALSE, message=FALSE}
#### Library and Data Imports ####

setwd('../')

## Load essential R packages
source("modules/Source.R")

## Load stored data (see `data` subdirectory)
source("modules/data_load.R")

## Create dataframes; perform desparity index calcuations; prep for plotting
source("modules/preprocessing.R")

```

# Final Project: Submission Links

*This should be the first section of your final project notebook. Fill out the following according to how you submitted your notebook!*

* github repository: https://github.com/TheRensselaerIDEA/COVID-Notebooks
* My github ID: *awi100*
* github issues addressed by this work: `#8` 
* Github branch name of my submitted notebook: *Imrie_final*
* link to merged notebook (post these to LMS!: 
    + https://github.com/TheRensselaerIDEA/COVID-Notebooks/blob/master/MATP-4400-FINAL/Final_Project_Andrew_Imrie.Rmd (example; Rmd version)
    + https://github.com/TheRensselaerIDEA/COVID-Notebooks/blob/master/MATP-4400-FINAL/Final_Project_Andrew_Imrie.html (example; HTML version)

# Overview & Problems Tackled

There has been so much talk in the news about unemployment and the numbers are staggering. I wanted to look at the trends from the last few years as well as the last few months to visualize the immediate impact COVID has had on unemployment. I also wanted to see if there was any strong correlation between the increase in the unemployment numbers and the number of covid cases in each state. Given the time and correct data sources, I would also like to dive deeper in New York state's data, and see if there are any specific groups where the economic impact is more severe. These potential groups include but are not limited to race, age, or gender. 

# Data Description

*Include data sources/locations, versions/dates, etc.* 

The first dataset I used is from the Bureau of Labor Statistics website. It is the monthly unemployment rate per state from 1976-present. It also included regional and national statistics. The data was only available as a txt file on the BLS website, so I downloaded it and wrote a Python script in order to clean the data. The cleaned data only has state data from 2018-present. In the Python script, I also created a dictionary of state codes and their respective states, and then added a state column to the filtered csv. I have attached screenshots of the Python script below.

(../pictures/uninsured_DI.png)


This data was pulled from the BLS website on April 28th 2020, and as such, it does not contain the unemployment data for April 2020. That data is set to be released on May 8th 2020, which is the due date of this project. The columns of the data breakdown as such:

*series_id*: series_id is a 20 length character code for what state the data came from as well as the type of data. All values begin with the 2 characters "LA" for Local Area. The following character is either a "S" or "U" for seasonally adjusted data or not. The next 15 characters maps the data to a particular state. For example, "ST0100000000000" is Alabama. The last 2 characters tell us what kind of unemployment data it is. For clarity purposes, I have broken down the 20 character series_id code in the following table:

| Represented         | Char Length | Possible Values                                                                                                                                                                                |
|---------------------|-------------|------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| Survey abbreviation | 2           | LA (Local Area)                                                                                                                                                                                |
| Seasonally          | 1           | S (Seasonal), U (Unadjusted)                                                                                                                                                                   |
| State               | 15          | ST0100000000000 (Alabama), ST0200000000000 (Alaska), etc.                                                                                                                                      |
| Measurement Code    | 2           | 03 (unemployment rate), 04 (unemployment), 05 (employment), 06 (labor force), 07 (employment-population ratio), 08 (labor force participation rate), 09 (civilian noninstitutional population) |

*year*: year is a numeric that simply represents the year in which the data is from. It ranges from 1976-2020.

*period*: period is a 3 length character code that represents the month the data is from. Values range from "M01" to "M12" and are based on the corresponded month of the year. "M13" is the states average for that year.

*value*: value is a numeric value that is the measured unemployment value. Because there are many different measurement types in the data set, the values for this column cannot be strictly defined. However, in this analysis we will be dealing with solely the unemployment rate data, which will simply be somewhere between 0.0 and about 25.0 (estimated max limit is based on the Great Depression maximum unemployment rate of 24.9).

The next data set I used is also from the Bureau of Labor Statistics website. It is simply the national unemployment rate by month. As with the previous dataset, I also had to write a Python script to clean this and convert it from .txt to .csv. 

# Results

*Break out your results by each problem you attacked*

## Problem 1 

*Describe the problem you are examining.  If there is background that is necessary for this problem, then put it here. Include any references.* 
The first problem that I am examining is the disparity index between state's unemployement rates and the US national average unemployment. Specifically, I am looking at maps of the months from Jan 2020 - March 2020. COVID-19 began to severely impact the United States in March, so this gives us two months of data just prior to when COVID-19 hit as well as a month of data from COVID-19 picking up. Personally, I believe the April 2020 data will tell a much more detailed story, but as previously mentioned, that data will only be available hours after the submission deadline.
 
### Methods

*How did you address the problem? What data did you use exactly? What methods did you use?* 
I addressed the problem by making a visualization of the disparity indexes amongst the states' unemployment rates vs. the national average. Before I could do this, I had to make a visualization of all the state's unemployment rates as well as the national average. After looking this visualization, I tested a few bin sizes and decided use a bin size of 0.2 in order to categorize each state. Once I did that, I used what COVIDMinder has been using for disparty index, which is ln(x/y). In this case, x will be the state's unemployment rate, and y will be the national average. 

	
### Results

```{r}
# Get unemployment data and name columns
Unemployment = read.csv('../data/csv/allstates_unemployment_filtered_s.csv', row.names = NULL)
colnames(Unemployment) = c("series_id", "state_id", "state", "year", "period", "value")

# Get state codes and name columns
state_codes = read.csv('../data/csv/states_unemployment_codes.csv', row.names = NULL)
colnames(state_codes) = c("state_id", "state")

national_avgs = read.csv('../data/csv/national_unemployment.csv', row.names = NULL)
colnames(national_avgs) = c("year", "month", "rate")

```

```{r}

# months = c("M01", "M02", "M03", "M04", "M05", "M06", "M07", "M08", "M09", "M10", "M11", "M12")
# years = c("2018", "2019", "2020")

jan_differences = c()
jan_disparity = c()

feb_differences = c()
feb_disparity = c()

march_differences = c()
march_disparity = c()

jan_2020_national <- national_avgs$rate[(national_avgs$year == "2020") & (national_avgs$month == "M01")]
jan_2020_national <- as.numeric((jan_2020_national))

feb_2020_national <- national_avgs$rate[(national_avgs$year == "2020") & (national_avgs$month == "M02")]
feb_2020_national <- as.numeric((feb_2020_national))

march_2020_national <- national_avgs$rate[(national_avgs$year == "2020") & (national_avgs$month == "M03")]
march_2020_national <- as.numeric((march_2020_national))

for (current_state in states$NAME){
  
  jan_2020_current_state <- Unemployment$value[Unemployment$state == current_state & Unemployment$year == "2020" & Unemployment$period == "M01"]
  jan_2020_current_state <- as.numeric(jan_2020_current_state)
  jan_diff <- jan_2020_national - jan_2020_current_state
  jan_disp <- log(jan_2020_national/jan_2020_current_state)
  jan_differences <- c(jan_differences,  jan_diff)
  jan_disparity <- c(jan_disparity, jan_disp)
  
  feb_2020_current_state <- Unemployment$value[Unemployment$state == current_state & Unemployment$year == "2020" & Unemployment$period == "M02"]
  feb_2020_current_state <- as.numeric(feb_2020_current_state)
  feb_diff <- feb_2020_national - feb_2020_current_state
  feb_disp <- log(feb_2020_national/feb_2020_current_state)
  feb_differences <- c(feb_differences, feb_diff)
  feb_disparity <- c(feb_disparity, feb_disp)
  
  march_2020_current_state <- Unemployment$value[Unemployment$state == current_state & Unemployment$year == "2020" & Unemployment$period == "M03"]
  march_2020_current_state <- as.numeric(march_2020_current_state)
  march_diff <- march_2020_national - march_2020_current_state
  march_disp <- log(march_2020_national/march_2020_current_state)
  march_differences <- c(march_differences,  march_diff)
  march_disparity <- c(march_disparity, march_disp)
}
```

```{r}

all_months <- cbind.data.frame(states$NAME, jan_disparity, feb_disparity, march_disparity)
colnames(all_months) = c("state", "Jan2020_ldi", "Feb2020_ldi", "Mar2020_ldi")
all_months$Jan2020_ldi <- as.numeric(all_months$Jan2020_ldi)
all_months$Feb2020_ldi <- as.numeric(all_months$Feb2020_ldi)
all_months$Mar2020_ldi <- as.numeric(all_months$Mar2020_ldi)

```


*What were the results on this problem?*
```{r echo = F, results = 'hide',  warning=FALSE, message=FALSE}
### Text used for captions and explanations

ldi_explanation_text <- "<h4><b>EXPLANATION OF VISUALIZATIONS:</b></h4> The goal of these visualizations is to examine 
the serverity of the economic impact COVID-19 has had across the states in terms of unemployment rates. There are three maps, 
which represent data from A common measure, the <i>disparity index</i> 
is used to represent thedifference between the observed unemployment rate in the state and the national unemployment rate.</p>
<p>The advantage of the disparity index is that represents how far off a target standard the observed rate is. </p>
<p>Mathematically,
<code>DI = log(x/y)</code> or <code>DI = log(y/x)</code>
depending upon whether being above or below the target is preferred. 
<ul>
<li>In this case, <i>x</i> will be some state's unemployment rate, and <i>y</i> would be the US national unemployment rate</li>
</ul>
</p>
<p>&nbsp;</p>"

### MEDIATION: Hospital Beds (USA)

us_unemployment_text <- "<h4><b>How does the unemployement rate across the United States compare to the national average?</b></h4>
<i>These maps compare the unemployment rate in US states vs the national unemployment rate 
For these three maps, <span style='color:#b2182b'><b>shades of red</b></span> indicate that a 
state's unemployment rate is lower than the national rate<br/><br>
Data sources: <br/><a href='https://download.bls.gov/pub/time.series/la/la.data.3.AllStatesS'>Bureau of Labor Statistics</a>"


jan_2020_text <- "<h2>January 2020</h4>"

```

 `r ldi_explanation_text`
 
 `r us_unemployment_text` 
 
 `r jan_2020_text` 

```{r}
    colors <- c("#253494","#4575B4", "#74ADD1","#ABD9E9","#f7f7f7","#FDAE61","#F46D43", "#D73027", "#BD0026")
    bins <- c(1, .8, .4, .2, .1, -.1, -.2, -.4, -.8, -1)
    pal2 <- leaflet::colorBin(colors, domain = all_months$Jan2020_ldi, bins = bins, reverse=TRUE)
    labels2 <- sprintf(
      "<strong>%s</strong><br/>
      National Unemployment Rate DI: %.2g",
      states$NAME, all_months$Jan2020_ldi
    ) %>% lapply(htmltools::HTML)
    
    leaflet(states.shapes) %>%
      setView(-96, 37.8, 4) %>% 
      addPolygons(
        fillColor = ~pal2(all_months$Jan2020_ldi),
        weight = 1,
        opacity = 1,
        color = "#330000",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels2,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~all_months$Jan2020_ldi, 
                opacity = 0.7, 
                title = "Disparity Index<br/>US vs State Unemployment Rate",
                position = "bottomright",
                labFormat = function(type, cuts, p) { n = length(cuts) 
                cuts[n] = paste0(cuts[n]," lower") 
                for (i in c(1,seq(2,(n-1)))){cuts[i] = paste0(cuts[i]," — ")} 
                cuts[2] = paste0(cuts[2]," higher") 
                paste0(str_remove(cuts[-n],"higher"), str_remove(cuts[-1],"—"))
                }
                ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light"))

```


```{r}
    colors <- c("#253494","#4575B4", "#74ADD1","#ABD9E9","#f7f7f7","#FDAE61","#F46D43", "#D73027", "#BD0026")
    bins <- c(1, .8, .4, .2, .1, -.1, -.2, -.4, -.8, -1)
    pal2 <- leaflet::colorBin(colors, domain = all_months$Feb2020_ldi, bins = bins, reverse=TRUE)
    labels2 <- sprintf(
      "<strong>%s</strong><br/>
      National Unemployment Rate DI: %.2g",
      states$NAME, all_months$Feb2020_ldi
    ) %>% lapply(htmltools::HTML)
    
    leaflet(states.shapes) %>%
      setView(-96, 37.8, 4) %>% 
      addPolygons(
        fillColor = ~pal2(all_months$Feb2020_ldi),
        weight = 1,
        opacity = 1,
        color = "#330000",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels2,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~all_months$Feb2020_ldi, 
                opacity = 0.7, 
                title = "Disparity Index<br/>US vs State Unemployment Rate",
                position = "bottomright",
                labFormat = function(type, cuts, p) { n = length(cuts) 
                cuts[n] = paste0(cuts[n]," lower") 
                for (i in c(1,seq(2,(n-1)))){cuts[i] = paste0(cuts[i]," — ")} 
                cuts[2] = paste0(cuts[2]," higher") 
                paste0(str_remove(cuts[-n],"higher"), str_remove(cuts[-1],"—"))
                }
                ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light"))

```




```{r}
    colors <- c("#253494","#4575B4", "#74ADD1","#ABD9E9","#f7f7f7","#FDAE61","#F46D43", "#D73027", "#BD0026")
    bins <- c(1, .8, .4, .2, .1, -.1, -.2, -.4, -.8, -1)
    pal2 <- leaflet::colorBin(colors, domain = all_months$Mar2020_ldi, bins = bins, reverse=TRUE)
    labels2 <- sprintf(
      "<strong>%s</strong><br/>
      National Unemployment Rate DI: %.2g",
      states$NAME, all_months$Mar2020_ldi
    ) %>% lapply(htmltools::HTML)
    
    leaflet(states.shapes) %>%
      setView(-96, 37.8, 4) %>% 
      addPolygons(
        fillColor = ~pal2(all_months$Mar2020_ldi),
        weight = 1,
        opacity = 1,
        color = "#330000",
        dashArray = "1",
        fillOpacity = 0.7,
        highlight = highlightOptions(
          weight = 5,
          color = "#666",
          dashArray = "",
          fillOpacity = 0.7,
          bringToFront = TRUE),
        label = labels2,
        labelOptions = labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto")) %>% 
      addLegend(pal = pal2, 
                values = ~all_months$Mar2020_ldi, 
                opacity = 0.7, 
                title = "Disparity Index<br/>US vs State Unemployment Rate",
                position = "bottomright",
                labFormat = function(type, cuts, p) { n = length(cuts) 
                cuts[n] = paste0(cuts[n]," lower") 
                for (i in c(1,seq(2,(n-1)))){cuts[i] = paste0(cuts[i]," — ")} 
                cuts[2] = paste0(cuts[2]," higher") 
                paste0(str_remove(cuts[-n],"higher"), str_remove(cuts[-1],"—"))
                }
                ) %>%
      addProviderTiles("MapBox", options = providerTileOptions(
        id = "mapbox.light"))

```





### Discussion

*Interpret results.  What were your findings?  What do they say about COVID-19?   What are the strengths and limitations of these results? Is there support for your findings from other sources? Include references as appropriate.*

## Problem 2
 
### Methods
	
### Results

### Discussion

*Do at least 2 problems. Add more as necessary *


# Summary and COVIDMINDER Recommendations

* Overall, what insights did you find  about the  COVID-19 epidemic in your analysis?    

* What recommendations do you have for COVIDMINDER for  Data utilization, Analytics, Visualizations, User interface design, etc.

    + Would you recommend that your analysis be included in COVIDMINDER?  Why or Why not?  
    + If not, is there additional work that might  improve the results?  Note that it is perfectly okay for you to recommend to not to include your analysis. Research doesn't always work out.  As a team we need to try a lot of different ideas in hopes of finding a few good ones to pursue.  *

*Think of yourself as a consultant reporting back on a particular aspect of the analysis and application design!*

# References

*Include any references to literature in support of your work*
*You might also wish to include references to unusual R packages essential to your work*

# Appendix

*Include here whatever you think is relevant to support the main content of your notebook. For example, you may have only include example figures above in your main text but include additional ones here* 


