---
title: "MATP-4400 COVID-19 Final Notebook"
author: "Tracy (Chuoxi) Chen"
date: "May 2020"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
# stick libraries here
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(maps)
library(cowplot)
library(tools)
```

# Final Project: Submission Links

* github repository: https://github.com/TheRensselaerIDEA/COVID-Notebooks
* My github ID: *Tracy29*
* github issues addressed by this work: `#6`
* Github branch name of my submitted notebook: *feature-6*
* link to merged notebook (post these to LMS!: 
    + https://github.com/TheRensselaerIDEA/COVID-Notebooks/blob/feature-6/MATP-4400-FINAL/Tracy_Chen_COVID_FINAL_2020.Rmd 
    + https://github.com/TheRensselaerIDEA/COVID-Notebooks/blob/feature-6/MATP-4400-FINAL/Tracy_Chen_COVID_FINAL_2020.pdf 

# Overview & Problems Tackled
In this project, I explored the connection between unhealthy behaviors and COVID 19 death for male and female. The approach was to first create visualizations of the disparity index for 3 common unhealthy behaviors: smoking, physical inactivity, and excessive drinking, across the 50 states for both male and female. Then I created a fourth visualization for the disparity index for COVID 19 death by states and gender. 

In the first problem, I investigated the disparity in smoking across the US. In the dataset, the percentage of male and female adult who reported smoking at least 100 cigarettes in their lifetime and currently smoke daily or some days in each state are recorded. From the disparity index, female adult in general tend to have lower percentage of smoker than male adult. In 

# Data Description

*Include data sources/locations, versions/dates, etc.* 

# Results

*Break out your results by each problem you attacked*

# Set up data
```{r }
# Read in health data
all_states.df <- read.csv("~/COVID-Notebooks/data/2019-Annual.csv")
all_states<- unique(all_states.df$State.Name) %>% sort()
# Sort data frame by states 
all_states.df<- all_states.df %>% 
  arrange(factor(all_states.df$State.Name, levels = all_states))

# America's Health Ranking (data source) defined health outcome as the interaction betweenthe 4 asepcts: behaviors, commuinty & environment, policy, and clinical care
# Pick data for behaviors as topic of this project
behaviors<- c("Excessive Drinking","Physical Inactivity", "Smoking")
behaviors.df <- filter(all_states.df, grepl(paste(behaviors, collapse="|"), Measure.Name))
```

```{r }
# Read in data about COVID 19 death
death_count<- read.csv("~/COVID-Notebooks/data/csv/Provisional_COVID-19_Death_Counts_by_Sex__Age__and_State.csv")

# Choose data that contains information about sex
pick<- c("Male Total", "Female Total")
death_count<- death_count %>%
  filter(grepl(paste(pick, collapse="|"), Sex)) %>%
  select(State, Sex, COVID.19.Deaths)

US_death<- death_count[death_count$State == "United States", ]
NYC_death<- death_count[death_count$State == "New York City", ]

# Remove irrelevant data
drop <- c("United States", "New York City", "Puerto Rico")
death_count<- death_count[!(death_count$State %in% drop), ]
row.names(death_count) <- NULL

# Add COVID 19 death information for NYC to NY's data for COVID 19 death
n<- which(grepl("New York", death_count$State))
death_count$COVID.19.Deaths[n[1]]<- death_count$COVID.19.Deaths[n[1]] + NYC_death$COVID.19.Deaths[1]
death_count$COVID.19.Deaths[n[2]]<- death_count$COVID.19.Deaths[n[2]] + NYC_death$COVID.19.Deaths[2]

# Calculate the average number of COVID 19 death for female and male
fmean_death<- US_death$COVID.19.Deaths[2]/53 # Divide by 53 because district of columbia and NYC counted is counted separately 
mmean_death<- US_death$COVID.19.Deaths[1]/53
US_death$mean_death<- c(fmean_death, mmean_death)

# Calculate the disparity index of each state's COVID 19 death count against the US value
death_count$death.disparity<- log(death_count$COVID.19.Deaths/US_death$mean_death)

# Reverse every two row from male, female to female, male to match the ordering for the health behavior data
death_count<- death_count[c(rbind(seq(2, nrow(death_count), 2), seq(1, nrow(death_count), 2))),]

# Categorize COVID 19 death count into 8 groups for female and male 
dcut_by<- c(5, 2, 1, .2, -.2, -1, -2, -5)
fdeath.df<- death_count %>% filter(grepl(paste("Female", collapse="|"), Sex))
fdeath.df$dgroup<- as.numeric(cut(fdeath.df$death.disparity, breaks = dcut_by))
mdeath.df<- death_count %>% filter(grepl(paste("Male", collapse="|"), Sex))
mdeath.df$dgroup<- as.numeric(cut(mdeath.df$death.disparity, breaks = dcut_by))
```

```{r }
sex<- c("Female", "Male")

# Filter out only the data that contain information with sex
behaviors_sex.df<- behaviors.df %>%
  filter(grepl(paste(sex, collapse="|"), Measure.Name)) %>%
  select(State.Name, Measure.Name, Value)

# Pick and remove the data pertaining to United States
US_behavior<- behaviors_sex.df[behaviors_sex.df$State.Name == "United States", ]
behaviors_sex.df<- behaviors_sex.df[behaviors_sex.df$State.Name != "United States", ]
```

## Problem 1 
smoking

*Describe the problem you are examining.  If there is background that is necessary for this problem, then put it here. Include any references.* 

 
### Methods

*How did you address the problem? What data did you use exactly? What methods did you use?* 
	
```{r, fig.width=10, fig.height=10}
# US_smoking value is the median value of states and DC
US_smoking<- US_behavior %>% filter(grepl(paste("Smoking", collapse="|"), Measure.Name))

# Filter out information about smoking 
smoking.df<- behaviors_sex.df %>%
  filter(grepl(paste("Smoking", collapse="|"), Measure.Name))
# Calculate the disparity index of each state's smoking percentage against the US value
smoking.df$value.disparity<- log(smoking.df$Value/US_smoking$Value)

# Separate smoking data by geneder
fsmoking.df<- smoking.df %>% 
  filter(grepl(paste("Female", collapse="|"), Measure.Name)) %>%
  # Add COVID 19 death disparity
  mutate(death.disparity= fdeath.df$death.disparity) %>% 
  mutate(dgroup= fdeath.df$dgroup)
fsmoking.df$region<- tolower(fsmoking.df$State.Name)

# Bin size for smoking disparity
vcut_by<- c(-1,-.5,-.25,-.1,0,.1,.25,.5,1)
# Categorize smoking percentage into 8 groups
fsmoking.df$vgroup<- as.numeric(cut(fsmoking.df$value.disparity, breaks = vcut_by))

msmoking.df<- smoking.df %>% 
  filter(grepl(paste("Male", collapse="|"), Measure.Name)) %>%
  mutate(death.disparity= mdeath.df$death.disparity) %>% 
  mutate(dgroup= mdeath.df$dgroup)
msmoking.df$region<- tolower(msmoking.df$State.Name)
msmoking.df$vgroup<- as.numeric(cut(msmoking.df$value.disparity, breaks = vcut_by))

# Access US map data from library(map)
us_states<- map_data("state")
# Abbreviate state names so the names can fit on the map
us_states$abbrev <- state.abb[match(toTitleCase(us_states$region), state.name)]

# Merge US map data and smoking data for female and male
us_fsmoking<- left_join(us_states, fsmoking.df)
us_msmoking<- left_join(us_states, msmoking.df)

# Use to locate state labels on the map
cnames <- aggregate(cbind(long, lat) ~ abbrev, data=us_fsmoking, 
                    FUN=function(x)mean(range(x)))

# Plot female and male smoking disparity 
vlabel<- c("-1 to -.5", "-.5 to -.25", "-.25 to -.1", "-.1 to 0", "0 to .1", ".1 to .25", ".25 to .5", ".5 to 1")
smoke_1<- ggplot(data = us_fsmoking, mapping = aes(x = long, y = lat, group = group, fill = as.factor(vgroup))) + 
  geom_polygon(color = "white", size = 0.2) + 
  with(cnames, annotate(geom="text", x = long, y=lat, label = abbrev, size = 2.5))+ coord_quickmap() + 
  theme_map()+labs(title = "US Female Smoking Disparity")+labs(fill = "Smoking Disparity") +
  scale_fill_brewer(palette = "RdYlBu", labels= vlabel, direction=-1)
smoke_2<- ggplot(data = us_msmoking, mapping = aes(x = long, y = lat, group = group, fill = as.factor(vgroup))) + 
  geom_polygon(color = "white", size = 0.2) + 
  with(cnames, annotate(geom="text", x = long, y=lat, label = abbrev, size = 2.5))+ coord_quickmap() + 
  theme_map()+labs(title = "US Male Smoking Disparity")+labs(fill = "Smoking Disparity") +
  scale_fill_brewer(palette = "RdYlBu", labels= vlabel, direction=-1)

# Combine the two maps
plot_grid(smoke_1, smoke_2, ncol = 1)

# Plot female and male COVID 19 death disparity
dlabel<- c("-5 to -2", "-2 to -1", "-1 to -.2", "-.2 to .2", ".2 to 1", "1 to 2", "2 to 5", "NA")
covid_1<- ggplot(data = us_fsmoking, mapping = aes(x = long, y = lat, group = group, fill = as.factor(dgroup))) + 
  geom_polygon(color = "white", size = 0.2) + 
  with(cnames, annotate(geom="text", x = long, y=lat, label = abbrev, size = 2.5))+ coord_quickmap() + 
  theme_map()+labs(title = "US Female COVID 19 Death Disparity")+labs(fill = "Death Disparity") +
  scale_fill_brewer(palette = "RdYlBu", labels= dlabel, direction=-1)
covid_2<- ggplot(data = us_msmoking, mapping = aes(x = long, y = lat, group = group, fill = as.factor(dgroup))) + 
  geom_polygon(color = "white", size = 0.2) + 
  with(cnames, annotate(geom="text", x = long, y=lat, label = abbrev, size = 2.5))+ coord_quickmap() + 
  theme_map()+labs(title = "US Male COVID 19 Death Disparity")+labs(fill = "Death Disparity") +
  scale_fill_brewer(palette = "RdYlBu", labels= dlabel, direction=-1)

# Combine the two maps
plot_grid(covid_1, covid_2, ncol = 1)
```

### Results

*What were the results on this problem?*


### Discussion

*Interpret results.  What were your findings?  What do they say about COVID-19?   What are the strengths and limitations of these results? Is there support for your findings from other sources? Include references as appropriate.*

## Problem 2
physical inactivity

### Methods
```{r, fig.width=10, fig.height=10}
# US_inactivity value is the median value of states and DC
US_inactivity<- US_behavior %>% filter(grepl(paste("Physical Inactivity", collapse="|"), Measure.Name))

# Filter out information about physical inactivity 
inactivity.df<- behaviors_sex.df %>%
  filter(grepl(paste("Physical Inactivity", collapse="|"), Measure.Name))
inactivity.df$value.disparity<- log(inactivity.df$Value/US_inactivity$Value)

# Separate physical inactivity data by geneder
finactivity.df<- inactivity.df %>% 
  filter(grepl(paste("Female", collapse="|"), Measure.Name))
finactivity.df$region<- tolower(finactivity.df$State.Name)
finactivity.df$vgroup<- as.numeric(cut(finactivity.df$value.disparity, breaks = vcut_by))

minactivity.df<- inactivity.df %>% 
  filter(grepl(paste("Male", collapse="|"), Measure.Name))
minactivity.df$region<- tolower(minactivity.df$State.Name)
minactivity.df$vgroup<- as.numeric(cut(minactivity.df$value.disparity, breaks = vcut_by))

# Merge US map data and physical inactivity data for female and male
us_finactivity<- left_join(us_states, finactivity.df)
us_minactivity<- left_join(us_states, minactivity.df)

# Plot femake and male physical inactivity disparity
inactivity_1<- ggplot(data = us_finactivity, mapping = aes(x = long, y = lat, group = group, fill = as.factor(vgroup))) + 
  geom_polygon(color = "white", size = 0.2) + 
  with(cnames, annotate(geom="text", x = long, y=lat, label = abbrev, size = 2.5))+ coord_quickmap() + 
  theme_map()+labs(title = "US Female Physical Inactivity Disparity")+labs(fill = "Physical Inactivity Disparity") +
  scale_fill_brewer(palette = "RdYlBu", labels= vlabel, direction=-1)
inactivity_2<- ggplot(data = us_minactivity, mapping = aes(x = long, y = lat, group = group, fill = as.factor(vgroup))) + 
  geom_polygon(color = "white", size = 0.2) + 
  with(cnames, annotate(geom="text", x = long, y=lat, label = abbrev, size = 2.5))+ coord_quickmap() + 
  theme_map()+labs(title = "US Male Physical Inactivity Disparity")+labs(fill = "Physical Inactivity Disparity") +
  scale_fill_brewer(palette = "RdYlBu", labels= vlabel, direction=-1)

# Combine the two maps
plot_grid(inactivity_1, inactivity_2, ncol = 1)
```
	
### Results

### Discussion

## Problem 3
From the data source, America's Health Ranking, it defined excessive drinking as "Adults who reported either binge drinking (having four or more [women] or five or more [men] drinks on one occasion in the past 30 days) or chronic drinking (having eight or more [women] or 15 or more [men] drinks per week)". There are many health risk associated with excessive drinking, such as the development of chronic disease and weakening immune system. I am interested in investigating the relationship between COVID-19 death and excessive drinking behaviors by comparing the assoiated value from America's Health Ranking on the percentage of people of years 18-45, 45-64, and 65+ and their corresponding age group's COVID-19 death by state. 

### Methods

```{r, fig.width=10, fig.height=10, warning = FALSE}
# US_drinking value is the median value of states and DC
US_drinking<- US_behavior %>% filter(grepl(paste("Excessive Drinking", collapse="|"), Measure.Name))

# Filter out information about excessive drinking
drinking.df<- behaviors_sex.df %>%
  filter(grepl(paste("Excessive Drinking", collapse="|"), Measure.Name))
drinking.df$value.disparity<- log(drinking.df$Value/US_drinking$Value)

# Separate excessive drinking data by geneder
fdrinking.df<- drinking.df %>% 
  filter(grepl(paste("Female", collapse="|"), Measure.Name))
fdrinking.df$region<- tolower(fdrinking.df$State.Name)
fdrinking.df$vgroup<- as.numeric(cut(fdrinking.df$value.disparity, breaks = vcut_by))

# Merge US map data and excessive drinking data for female and male
mdrinking.df<- drinking.df %>% 
  filter(grepl(paste("Male", collapse="|"), Measure.Name))
mdrinking.df$region<- tolower(mdrinking.df$State.Name)
mdrinking.df$vgroup<- as.numeric(cut(mdrinking.df$value.disparity, breaks = vcut_by))

# Merge US map data and excessive drinking data for female and male
us_fdrinking<- left_join(us_states, fdrinking.df)
us_mdrinking<- left_join(us_states, mdrinking.df)

# Plot femake and male excessive drinking disparity
drinking_1<- ggplot(data = us_fdrinking, mapping = aes(x = long, y = lat, group = group, fill = as.factor(vgroup))) + 
  geom_polygon(color = "white", size = 0.2) + 
  with(cnames, annotate(geom="text", x = long, y=lat, label = abbrev, size = 2.5))+ coord_quickmap() + 
  theme_map()+labs(title = "US Female Excessive Drinking Disparity")+labs(fill = "Excessive Drinking Disparity") +
  scale_fill_brewer(palette = "RdYlBu", labels= vlabel, direction=-1)
drinking_2<- ggplot(data = us_mdrinking, mapping = aes(x = long, y = lat, group = group, fill = as.factor(vgroup))) + 
  geom_polygon(color = "white", size = 0.2) + 
  with(cnames, annotate(geom="text", x = long, y=lat, label = abbrev, size = 2.5))+ coord_quickmap() + 
  theme_map()+labs(title = "US Male Excessive Drinking Disparity")+labs(fill = "Excessive Drinking Disparity") +
  scale_fill_brewer(palette = "RdYlBu", labels= vlabel, direction=-1)

# Combine the two maps
plot_grid(drinking_1, drinking_2, ncol = 1)
```

### Results

### Discussion

# Summary and COVIDMINDER Recommendations

* Overall, what insights did you find  about the  COVID-19 epidemic in your analysis?    

* What recommendations do you have for COVIDMINDER for  Data utilization, Analytics, Visualizations, User interface design, etc.

    + Would you recommend that your analysis be included in COVIDMINDER?  Why or Why not?  
    + If not, is there additional work that might  improve the results?  Note that it is perfectly okay for you to recommend to not to include your analysis. Research doesn't always work out.  As a team we need to try a lot of different ideas in hopes of finding a few good ones to pursue.  *

*Think of yourself as a consultant reporting back on a particular aspect of the analysis and application design!*

# References

*Include any references to literature in support of your work*
*You might also wish to include references to unusual R packages essential to your work*

# Appendix

*Include here whatever you think is relevant to support the main content of your notebook. For example, you may have only include example figures above in your main text but include additional ones here* 


